---
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    includes:
      in_header: "../preamble.tex"
      before-body: "../before-body.tex"
    keep_tex: true
    number_sections: false
    toc: false
papersize: a4
classoption: "twoside, 11pt"
geometry: "top=2cm, outer=3.5cm, bottom=3cm, inner=3.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, fig.align = "center", fig.pos = 'tb')

library("assertthat")
library("fable")
library("feasts")
library("ggtext")
library("janitor")
library("lubridate")
library("magick")
library("scales")
library("sf")
library(SpatialKDE)
library(spdep)
library("tsibble")
library("tidyverse")
```



```{r functions}
source("https://raw.githubusercontent.com/mpjashby/r-helper/main/helpers.R")

# define colour scheme, based on the UCL colour scheme defined at
# https://www.ucl.ac.uk/cam/brand/guidelines/colour
ucl_colours <- tribble(
  ~name, ~hex_code,
  "Dark Green", "#555025",
  "Dark Red", "#651D32",
  "Dark Purple", "#4B384C",
  "Dark Blue", "#003D4C",
  "Dark Brown", "#4E3629",
  "Mid Green", "#8F993E",
  "Mid Red", "#93272C",
  "Mid Purple", "#500778",
  "Mid Blue", "#002855",
  "Stone", "#D6D2C4",
  "Bright Green", "#B5BD00",
  "Bright Red", "#D50032",
  "Bright Blue", "#0097A9",
  "Bright Pink", "#AC145A",
  "Light Green", "#BBC592",
  "Light Red", "#E03C31",
  "Light Purple", "#C6B0BC",
  "Light Blue", "#8DB9CA",
  "Yellow", "#F6BE00",
  "Orange", "#EA7600",
  "Grey", "#8C8279",
  "Blue Celeste", "#A4DBE8"
)

ucl_colours_list <- set_names(ucl_colours$hex_code, ucl_colours$name)

search_type_colours <- c(
  "drugs" = ucl_colours_list[["Bright Green"]], 
  "weapons" = ucl_colours_list[["Light Purple"]], 
  "stolen goods" = ucl_colours_list[["Bright Blue"]], 
  "firearms" = ucl_colours_list[["Bright Red"]], 
  "other" = ucl_colours_list[["Blue Celeste"]], 
  "unknown" = ucl_colours_list[["Grey"]],
  "weapons (based on reasonable suspicion)" = ucl_colours_list[["Bright Pink"]],
  "weapons (based on authorisation)" = ucl_colours_list[["Light Red"]]
)

to_text <- partial(knitr::combine_words, oxford_comma = FALSE)

prop_of_total <- function (data, var, criteria) {
  stops %>% 
    dplyr::filter(date >= date_start) %>% 
    dplyr::count({{var}}) %>% 
    janitor::adorn_percentages(denominator = "col") %>% 
    dplyr::filter({{criteria}}) %>% 
    purrr::pluck("n")
}

mid_point <- function (a, b) {
  if (a > b) {
    b + ((a - b) / 2)
  } else {
    a + ((b - a) / 2)
  }
}

describe_quarter <- function (date) {
  stringr::str_glue(
    dplyr::recode(
      ceiling(lubridate::month(date) / 3), 
      `1` = "first", 
      `2` = "second", 
      `3` = "third", 
      `4` = "fourth"
    ),
    " quarter of ",
    lubridate::year(date)
  )
}

theme_stop_search <- function (...) {
  theme_minimal(...) %+replace%
    theme(
      axis.ticks = element_line(colour = "grey92"),
      axis.title = element_text(size = 9, hjust = 1),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(6, "mm"),
      legend.position = "bottom",
      legend.spacing.x = unit(2, "mm"),
      legend.title = element_text(size = 9),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.caption = element_text(size = 9, colour = "grey33", hjust = 1, 
                                  margin = margin(t = 3)),
      plot.margin = margin(t = 0, b = 9),
      plot.tag = element_text(size = 12, face = "bold", colour = "grey33", 
                              hjust = 0),
      plot.tag.position = c(0.01, 0.01),
      plot.title = element_text(face = "plain", size = 12, hjust = 0, 
                                margin = margin(t = 12, b = 6)),
      plot.title.position = "plot",
      plot.subtitle = element_text(face = "bold", size = 14, hjust = 0, 
                                margin = margin(b = 9)),
      strip.text.y = element_text(angle = 0, hjust = 0)
    )
}

# add logo to chart
add_logo <- function (chart, title, title_as_subtitle = TRUE, set_caption = TRUE) {

  scs_logo <- here::here("output/ucl-banner-land-yellow-rgb.png") %>% 
    png::readPNG() %>% 
    grid::rasterGrob(
      x = unit(0, "npc"), 
      y = unit(1, "npc"), 
      width = unit(1, "npc"),
      hjust = 0,
      vjust = 1
    )
  
  if (title_as_subtitle) {
    chart <- chart + labs(
      title = "Stop and search in London",
      subtitle = title
    )
  }
  
  ggpubr::ggarrange(
    grid::grobTree(
      scs_logo,
      grid::textGrob(
        "INSTITUTE FOR GLOBAL CITY POLICING",
        x = unit(1, "lines"),
        y = unit(1, "npc") - unit(1.2, "lines"),
        hjust = 0,
        vjust = 1,
        gp = grid::gpar(fontfamily = "Helvetica", fontsize = 7, 
                        fontface = "bold")
      )
    ),
    ggplotGrob(chart),
    ggpubr::ggarrange(
      grid::textGrob(
        str_glue(
          "Data: Home Office, {lubridate::year(Sys.time())} | Licence: ",
          "Creative Commons Attribution | Author: Dr Matt Ashby"
        ),
        x = unit(0, "npc"), 
        hjust = 0,
        gp = grid::gpar(col = "grey33", fontfamily = "Helvetica", fontsize = 8, 
                        lineheight = 1)
      ),
      grid::textGrob(
        ifelse(set_caption, "More details: bit.ly/stop-search-london-2020", ""),
        x = unit(1, "npc"), 
        hjust = 1,
        gp = grid::gpar(col = "grey33", fontfamily = "Helvetica", 
                        fontsize = 8, fontface = "bold", lineheight = 1)
      ),
      ncol = 2,
      widths = c(3, 1)
    ),
    ncol = 1,
    hjust = 0, 
    heights = c(3, 25.5, 1)
  )
  
}

# save chart
save_chart <- function (chart, slug, title, height = 400, width = 600, ...) {
  ggsave(
    filename = str_glue("{report_dir}/{report_q}_{slug}.png"), 
    plot = add_logo(chart, title, ...),
    device = "png", 
    width = width / 72, 
    height = height / 72, 
    units = "in",
    bg = "white"
  )
}
```



```{r load stops data, include=FALSE}
stops <- here::here("analysis-data/stops.rds") %>% 
  read_rds()
```




```{r create report date strings}
date_end <- stops %>% 
  pluck("date") %>% 
  max() %>% 
  as_date() %>% 
  ceiling_date(unit = "months") - days(1)
date_start <- date_end - years(1) + days(1)

report_period <- str_glue(
  format(date_start, ifelse(year(date_start) == year(date_end), "%B", "%B %Y")),
  " to ",
  format(date_end, "%B %Y")
)

report_between <- str_replace(report_period, " to ", " and ")
report_span <- str_replace(report_period, " to ", "–")
# `report_q` is used in file names -- it refers to quarters rather than years
# because the reports are produced quarterly
report_q <- str_glue(
  year(date_end),
  "_",
  case_when(
    month(date_end) %in% 1:3 ~ "q1",
    month(date_end) %in% 4:6 ~ "q2",
    month(date_end) %in% 7:9 ~ "q3",
    month(date_end) %in% 10:12 ~ "q4",
    TRUE ~ NA_character_
  )
)

# Create directory for files from this report
report_dir <- here::here(str_glue("output/{report_q}"))
if (!dir.exists(report_dir)) dir.create(report_dir)
```



```{r load other data, include=FALSE}
historical_stops <- here::here("analysis-data/historical-stop-data.rds") %>% 
  read_rds() %>% 
  count(financial_year, wt = n) %>% 
  filter(financial_year < ymd("2018-04-01"))

people <- read_rds(here::here("analysis-data/people.rds"))

imd <- read_rds(here::here("analysis-data/imd.rds"))

london_stations <- here::here("analysis-data/stations.gpkg") %>% 
  st_read(quiet = TRUE) %>% 
  st_transform(27700) %>% 
  rownames_to_column(var = "id") %>% 
  mutate(id = as.numeric(id)) %>% identity()

# Download data if needed
tribble(
  ~file, ~url,
  "lsoa", "https://opendata.arcgis.com/datasets/8bbadffa6ddc493a94078c195a1e293b_0.geojson",
  "borough", "https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson",
  "ward_lookup", "https://opendata.arcgis.com/datasets/e169bb50944747cd83dcfb4dd66555b1_0.csv",
  "ward", "https://opendata.arcgis.com/datasets/446d684e667a4d91bd54a073efff207d_0.geojson"
) %>% 
  pwalk(function (url, file) {
    file_type <- str_extract(url, "\\.\\w+$")
    file_name <- str_glue("{report_dir}/data_{file}{file_type}")
    if (!file.exists(file_name)) {
      download.file(url = url, destfile = file_name, timeout = 600)
    }
  })

# https://geoportal.statistics.gov.uk/datasets/lower-layer-super-output-areas-december-2011-boundaries-generalised-clipped-bgc-ew-v3
lsoa <- str_glue("{report_dir}/data_lsoa.geojson") %>% 
  read_sf() %>% 
  select(lsoa_code = LSOA11CD, lsoa_name = LSOA11NM, geometry)

# https://geoportal.statistics.gov.uk/datasets/local-authority-districts-december-2019-boundaries-uk-bgc
london_boroughs <- str_glue("{report_dir}/data_borough.geojson") %>% 
  read_sf() %>%
  clean_names() %>% 
  filter(str_detect(lad19cd, "^E09")) %>% 
  rename(borough = lad19nm) %>% 
  st_transform(27700)

london_outline <- london_boroughs %>% 
  st_union() %>% 
  st_sf() %>% 
  mutate(region = "London")

ward_lad_lookup <- str_glue("{report_dir}/data_ward_lookup.csv") %>% 
  read_csv() %>% 
  clean_names() %>% 
  filter(str_detect(lad19cd, "^E09")) %>% 
  select(wd19cd, borough = lad19nm)

london_wards <- str_glue("{report_dir}/data_ward.geojson") %>% 
  read_sf() %>% 
  clean_names() %>% 
  st_transform(27700) %>% 
  left_join(ward_lad_lookup, by = "wd19cd") %>% 
  filter(!is.na(borough)) %>% 
  select(ward = wd19nm, borough, geometry)
```



```{r check data values are as expected, include=FALSE}
# check data only contains known variables
assert_that(setequal(
  names(stops), 
  c(
    "force",
    "type",
    "date",
    "latitude",
    "longitude",
    "gender",
    "age_range",
    "self_defined_ethnicity",
    "officer_defined_ethnicity",
    "legislation",
    "object_of_search",
    "outcome"
  )
))

# check factors only have known levels
assert_that(all(unique(stops$force) %in% c("British Transport Police", "City of London Police", "Metropolitan Police Service")))
assert_that(all(unique(stops$type) %in% c("Person search", "Person and Vehicle search", "Vehicle search")))
assert_that(all(unique(stops$gender) %in% c("male", "female", "other", NA_character_)))
assert_that(all(unique(stops$age_range) %in% c("under 10", "10-17", "18-24", "25-34", "over 34", NA_character_)))
assert_that(all(unique(stops$self_defined_ethnicity) %in% c(
  "Asian/Asian British - Any other Asian background", 
  "Asian/Asian British - Bangladeshi", 
  "Asian/Asian British - Chinese", 
  "Asian/Asian British - Indian", 
  "Asian/Asian British - Pakistani", 
  "Black/African/Caribbean/Black British - African", 
  "Black/African/Caribbean/Black British - Any other Black/African/Caribbean background",
  "Black/African/Caribbean/Black British - Caribbean", 
  "Mixed/Multiple ethnic groups - Any other Mixed/Multiple ethnic background", 
  "Mixed/Multiple ethnic groups - White and Asian", 
  "Mixed/Multiple ethnic groups - White and Black African", 
  "Mixed/Multiple ethnic groups - White and Black Caribbean", 
  "Other ethnic group - Any other ethnic group", 
  "Other ethnic group - Not stated", 
  "White - Any other White background", 
  "White - English/Welsh/Scottish/Northern Irish/British", 
  "White - Irish",
  NA_character_
)))
assert_that(all(unique(stops$officer_defined_ethnicity) %in% c("Asian", "Black", "Other", "White", NA_character_)))
assert_that(all(unique(stops$legislation) %in% c(
  "Criminal Justice Act 1988 (section 139B)",
  "Criminal Justice and Public Order Act 1994 (section 60)", 
  "Deer Act 1991 (section 12)", 
  "Firearms Act 1968 (section 47)", 
  "Misuse of Drugs Act 1971 (section 23)", 
  "Police and Criminal Evidence Act 1984 (section 1)", 
  "Police and Criminal Evidence Act 1984 (section 6)", 
  "Psychoactive Substances Act 2016 (s36(2))", 
  "Wildlife and Countryside Act 1981 (section 19)",
  NA_character_
)))
assert_that(all(unique(stops$object_of_search) %in% c(
  "anything to threaten or harm anyone", 
  "article for use in theft", 
  "articles for use in criminal damage",
  "controlled drugs", 
  "crossbows", 
  "evidence of offences under the act", 
  "evidence of wildlife offences",
  "firearms", 
  "fireworks",
  "goods on which duty has not been paid etc.", 
  "offensive weapons",
  "psychoactive substances", 
  "stolen goods",
  NA_character_
)))
assert_that(all(unique(stops$outcome) %in% c(
  "a no further action disposal", 
  "arrest",
  "caution (simple or conditional)",
  "community resolution",
  "khat or cannabis warning",
  "local resolution",
  "nothing found - no further action",
  "offender given drugs possession warning",
  "offender given penalty notice",
  "penalty notice for disorder",
  "summons / charged by post", 
  "suspect arrested",
  "suspect summonsed to court",
  NA_character_
)))

# check date/time is POSIXct
assert_that(all(is.POSIXct(stops$date)))

# check co-ordinates are valid
# bounding box co-ordinates from https://gist.github.com/graydon/11198540
assert_that(all(is.double(stops$longitude)))
assert_that(all(stops$longitude >= -180, na.rm = TRUE))
assert_that(all(stops$longitude <= 180, na.rm = TRUE))
assert_that(all(is.double(stops$latitude)))
assert_that(all(stops$latitude >= -90, na.rm = TRUE))
assert_that(all(stops$latitude <= 90, na.rm = TRUE))
```



```{r remove non-London BTP stops}
# BTP data includes all stops, wherever in England and Wales they occurred. We
# can remove those stops occurring outside the Greater London boundary, although
# this means that BTP stops without a location (some of which are likely to have
# occurred in London) will also be excluded.
btp_london_stops <- stops %>% 
  filter(force == "British Transport Police", !is.na(latitude), 
         !is.na(longitude)) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE) %>% 
  st_transform(27700) %>% 
  st_join(london_outline) %>% 
  filter(region == "London") %>% 
  as_tibble() %>% 
  select(-region)

stops <- stops %>% 
  filter(force != "British Transport Police") %>% 
  bind_rows(btp_london_stops)
```



```{r harmonise variables}
stops <- stops %>% 
  mutate(
    object_of_search = case_when(
      legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
        "offensive weapons",
      TRUE ~ object_of_search
    ),
    object = recode(
      object_of_search,
      "controlled drugs" = "drugs",
      "offensive weapons" = "weapons",
      "evidence of offences under the act" = "unknown",
      "articles for use in criminal damage" = "other",
      "article for use in theft" = "other",
      "evidence of wildlife offences" = "other",
      "fireworks" = "other",
      "goods on which duty has not been paid etc." = "other",
      "anything to threaten or harm anyone" = "weapons",
      "crossbows" = "weapons",
      "psychoactive substances" = "drugs"
    ),
    outcome = recode(
      outcome,
      "suspect arrested" = "arrest",
      "community resolution" = "community/local resolution",
      "local resolution" = "community/local resolution",
      "caution (simple or conditional)" = "caution",
      "offender cautioned" = "caution",
      "khat or cannabis warning" = "drugs warning",
      "offender given penalty notice" = "fixed penalty",
      "penalty notice for disorder" = "fixed penalty",
      "offender given drugs possession warning" = "drugs warning",
      "a no further action disposal" = "no further action",
      "nothing found - no further action" = "no further action",
      "suspected psychoactive substances seized - no further action" = 
        "no further action",
      "summons / charged by post" = "charged by post",
      "suspect summonsed to court" = "charged by post"
    ),
    year = year(date),
    year_month = yearmonth(date)
  ) %>% 
  replace_na(list(object = "unknown", outcome = "unknown")) %>% 
  mutate(
    outcome = fct_relevel(
      outcome,
      "arrest",
      "charged by post",
      "caution",
      "fixed penalty",
      "community/local resolution",
      "drugs warning",
      "no further action",
      "unknown"
    )
  ) %>% 
  filter(year >= 2018)
```



```{r count stops and analyse trends, include=FALSE}
monthly_counts <- stops %>% 
  count(year_month) %>% 
  mutate(
    year_month = as_date(year_month),
    diff = n - lag(n),
    diff_perc = diff / lag(n)
  )

monthly_counts_feats <- monthly_counts %>% 
  filter(as_date(year_month) >= date_start) %>% 
  pluck("n") %>% 
  feasts::feat_stl(.period = 12) %>% 
  as.list()

monthly_counts_anomaly <- timetk::tk_anomaly_diagnostics(
  monthly_counts, 
  .date_var = year_month, 
  .value = n
)

total_searches <- stops %>% 
  filter(date >= date_start) %>% 
  count() %>% 
  pluck("n", 1)

previous_total_searches <- stops %>% 
  filter(date >= date_start - months(3), date <= date_end - months(3)) %>% 
  count() %>% 
  pluck("n", 1)

total_searches_change <- as.character(cut(
  (total_searches - previous_total_searches) / previous_total_searches,
  breaks = c(-Inf, -0.05, -0.01, 0.01, 0.05, Inf),
  labels = c(
    "falling to",
    "falling slightly to",
    "staying roughly constant at",
    "increasing slightly to",
    "increasing to"
  )
))

prop_weapons_model <- stops %>% 
  filter(date >= date_start) %>% 
  count(year_month, weapons = object == "weapons") %>% 
  group_by(year_month) %>% 
  mutate(prop_weapons = n / sum(n)) %>% 
  ungroup() %>% 
  filter(weapons) %>% 
  as_tsibble(index = year_month) %>%
  model(model = TSLM(prop_weapons ~ trend())) %>%
  broom::tidy()
```



```{r search types}
main_types <- c("drugs", "firearms", "stolen goods", "weapons")

searches_by_type <- count(stops, year_month, object)
```



```{r search results by month}
search_result_annual_counts <- stops %>% 
  filter(date >= date_start, outcome != "unknown") %>% 
  mutate(positive = outcome != "no further action") %>% 
  count(positive) %>% 
  mutate(prop = n / sum(n))

search_result_month_counts <- stops %>% 
  mutate(
    positive = !outcome %in% c("no further action", "unknown"),
    year_month = as_date(year_month)
  ) %>% 
  count(year_month, positive) %>% 
  group_by(year_month) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()
```



```{r search disparity}
person_stops <- filter(
  stops, 
  type %in% c("Person search", "Person and Vehicle search")
)

age_counts <- person_stops %>% 
  filter(date >= date_start, !is.na(age_range)) %>% 
  tabyl(age_range)

sde_counts <- person_stops %>% 
  filter(date >= date_start, !is.na(self_defined_ethnicity)) %>% 
  mutate(
    sde_group = case_when(
      self_defined_ethnicity == "Other ethnic group - Not stated" ~ 
        NA_character_,
      str_detect(self_defined_ethnicity, "^Asian") ~ "Asian/Asian British",
      str_detect(self_defined_ethnicity, "^Black") ~ "Black/Black British",
      str_detect(self_defined_ethnicity, "^Mixed") ~ 
        "from multiple ethnic groups",
      str_detect(self_defined_ethnicity, "^Other") ~ 
        "from another ethnic group",
      str_detect(self_defined_ethnicity, "^White") ~ "white",
      TRUE ~ NA_character_
    )
  ) %>% 
  tabyl(sde_group) %>% 
  arrange(desc(valid_percent))

stops_disparity <- person_stops %>% 
  filter(
    date >= date_start, 
    !is.na(gender),
    gender %in% c("female", "male"),
    !is.na(age_range), 
    age_range != "under 10",
    !str_detect(self_defined_ethnicity, "^Other ethnic group")
  ) %>% 
  rename(sex = gender) %>% 
  count(sex, age_range, self_defined_ethnicity, name = "searches") %>% 
  left_join(people, by = c("sex", "age_range", "self_defined_ethnicity")) %>% 
  separate(self_defined_ethnicity, c("ethnic_group", "ethnicity"), " - ") %>% 
  group_by(age_range, ethnic_group, sex) %>% 
  summarise(people = sum(people), searches = sum(searches), 
            .groups = "drop") %>% 
  mutate(
    ethnic_group = str_extract(ethnic_group, "^\\w+"),
    age_label = case_when(
      age_range %in% c("under 10", "10-17") & sex == "female" ~ "girls",
      age_range %in% c("under 10", "10-17") & sex == "male" ~ "boys",
      !age_range %in% c("under 10", "10-17") & sex == "female" ~ "women",
      !age_range %in% c("under 10", "10-17") & sex == "male" ~ "men",
    ),
    group = str_glue("**{age_label}** aged **{age_range}** identifying as ",
                     "**{ethnic_group}**"),
    search_rate = searches / (people / 1000)
  )

stops_disparity_object <- person_stops %>% 
  mutate(
    object = case_when(
      object == "weapons" & 
        legislation == "Police and Criminal Evidence Act 1984 (section 1)" ~ 
        "weapons (based on reasonable suspicion)",
      object == "weapons" & 
        legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
        "weapons (based on authorisation)",
      object == "weapons" ~ "unknown",
      TRUE ~ object
    )
  ) %>% 
  filter(
    date >= date_start, 
    !is.na(age_range), 
    age_range != "under 10",
    !is.na(gender),
    gender %in% c("female", "male"),
    !object %in% c("other", "unknown"),
    !str_detect(self_defined_ethnicity, "^Other ethnic group")
  ) %>% 
  select(object, sex = gender, age_range, self_defined_ethnicity) %>% 
  mutate_if(is.character, as.factor) %>% 
  count(object, sex, age_range, self_defined_ethnicity, name = "searches", .drop = FALSE) %>% 
  left_join(people, by = c("sex", "age_range", "self_defined_ethnicity")) %>% 
  separate(self_defined_ethnicity, c("ethnic_group", "ethnicity"), " - ") %>% 
  group_by(object, age_range, ethnic_group, sex) %>% 
  summarise(people = sum(people), searches = sum(searches), 
            .groups = "drop") %>% 
  mutate(
    ethnic_group = str_extract(ethnic_group, "^\\w+"),
    age_label = case_when(
      age_range %in% c("under 10", "10-17") & sex == "female" ~ "girls",
      age_range %in% c("under 10", "10-17") & sex == "male" ~ "boys",
      !age_range %in% c("under 10", "10-17") & sex == "female" ~ "women",
      !age_range %in% c("under 10", "10-17") & sex == "male" ~ "men",
    ),
    group = str_glue("{age_label}, {age_range}, {ethnic_group}"),
    search_rate = searches / (people / 1000)
  )

stops_model <- stops_disparity %>% 
  mutate(
    age_range = fct_relevel(age_range, "over 34"),
    ethnic_group = fct_relevel(ethnic_group, "White"),
    sex = fct_relevel(sex, "female")
  ) %>% 
  {
    glm(log(search_rate) ~ sex + age_range + ethnic_group, data = .)
  } %>% 
  broom::tidy() %>% 
  arrange(desc(estimate)) %>% 
  # keep only model terms that are greater than zero, significant and greater 
  # than the largest co-efficients for ethnicity
  filter(
    estimate > 0,
    estimate > 
      pluck(filter(., str_detect(term, "^ethnic_group")), "estimate", 1),
    p.value < 0.05
  ) %>%
  mutate(
    term = case_when(
      str_detect(term, "^age_range") ~
        str_glue("being aged {str_remove(term, '^age_range')}"),
      str_detect(term, "^sex") ~ str_glue("being {str_remove(term, '^sex')}"),
      TRUE ~ term
    )
  ) %>% 
  pull("term") %>% 
  {
    if ("being aged 10-17" %in% . & "being aged 18-24" %in% . & 
        "being aged 25-34" %in% .) {
      c(str_subset(., "^being aged", negate = TRUE), "being aged under 35")
    } else if ("being aged 10-17" %in% . & "being aged 18-24" %in% .) {
      c(str_subset(., "^being aged", negate = TRUE), "being aged under 25")
    } else if ("being aged 18-24" %in% . & "being aged 25-34" %in% .) {
      c(str_subset(., "^being aged", negate = TRUE), "being aged 18-34")
    } else {
      .
    }
  }

overall_search_rate <- sum(stops_disparity$searches) / (sum(stops_disparity$people) / 1000)
```



```{r searches by borough and ward, include=FALSE}
searches_by_borough <- stops %>% 
  filter(!is.na(latitude), !is.na(longitude), date >= date_start) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(27700) %>% 
  st_join(london_boroughs) %>% 
  as_tibble() %>% 
  count(borough, name = "searches") %>% 
  filter(!is.na(borough)) %>% 
  mutate(searches_per_month = searches / 12) %>% 
  arrange(desc(searches))

searches_by_ward <- stops %>% 
  filter(!is.na(latitude), !is.na(longitude), date >= date_start) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(27700) %>% 
  st_join(london_wards) %>% 
  as_tibble() %>% 
  count(borough, ward, name = "searches") %>% 
  filter(!is.na(borough), !is.na(ward)) %>% 
  arrange(desc(searches))

stops_by_lsoa <- stops %>% 
  filter(!is.na(latitude), !is.na(longitude), date >= date_start) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_join(lsoa) %>% 
  st_set_geometry(NULL) %>% 
  mutate(object = case_when(
    object == "weapons" & 
      legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
      "weapons under section 60",
    object == "weapons" ~ "weapons under section 1",
    TRUE ~ object
  ))

# this code first identifies all the LSOAs in London, then counts the searches 
# in each one, since this keeps the LSOAs with zero searches
stop_concentration <- lsoa %>% 
  st_transform(27700) %>% 
  st_join(london_boroughs) %>% 
  filter(!is.na(borough)) %>% 
  st_set_geometry(NULL) %>% 
  select(lsoa_code) %>% 
  left_join(
    count(stops_by_lsoa, lsoa_code, name = "searches"), 
    by = "lsoa_code"
  ) %>% 
  replace_na(list(searches = 0)) %>% 
  arrange(desc(searches)) %>% 
  mutate(
    csum_searches = cumsum(searches),
    csum_searches = csum_searches / max(csum_searches),
    perc_rank = row_number() / n()
  ) %>% 
  filter(csum_searches > 0.5) %>% 
  slice(1) %>% 
  pull(perc_rank)

low_search_boroughs <- searches_by_borough %>% 
  filter(searches < searches_by_ward$searches[1]) %>% 
  pull(borough)
```



```{r headlines}
headline1 <- str_glue(
  "Police in London stopped and searched ",
  comma(pluck(count(filter(stops, date >= date_start)), "n", 1)),
  " people and vehicles in the 12~months from {report_period}. The number ",
  "of searches has generally ",
  ifelse(monthly_counts_feats$linearity > 0, "increased", "decreased"),
  " over the past year."
)

headline2 <- str_replace_all(str_glue(
  percent(pluck(arrange(janitor::adorn_percentages(
    count(filter(stops, date >= date_start), object), 
  denominator = "col"), desc(n)), "n", 1)),
  " of searches in that period were for ",
  pluck(arrange(janitor::adorn_percentages(
    count(filter(stops, date >= date_start), object), 
  denominator = "col"), desc(n)), "object", 1),
  ", with ",
  percent(pluck(filter(search_result_annual_counts, !positive), "prop")),
  " of all searches resulting in no further action."
), "%", "\\\\%")

# headline3 <- str_glue(
#   "Different demographic groups are searched at different rates, with ",
#   str_remove_all(pluck(arrange(stops_disparity, desc(search_rate)), "group", 1), 
#                  "\\*\\*"),
#   " being ", 
#   number(
#     pluck(arrange(stops_disparity, desc(search_rate)), "search_rate", 1) / 
#       overall_search_rate
#   ), 
#   " times more likely to be searched than the population as a whole."
# )

headline4 <- str_replace_all(str_glue(
  "Searches are heavily concentrated in some areas -- half of all searches ",
  "occurred in {percent(stop_concentration)} of neighbourhoods."
), "%", "\\\\%")
```



<!-- single space at the end of each sentence -->
\frenchspacing

<!-- left align as per UCL style guide -->
\raggedright

<!-- prevent vertical justification of paragraphs -->
\raggedbottom

<!-- The following blocks of Latex code produce the title page -->

\begin{textblock*}{21cm}(0mm, 0mm)
\includegraphics[width=21cm,height=29.7cm]{`r str_glue("../{report_q}/cover_image_{report_q}.jpg")`}
\end{textblock*}

\begin{textblock*}{21cm}(0mm, 0mm)
\includegraphics[width=21cm]{../ucl-banner-port-yellow-rgb-lg.png}
\end{textblock*}

\begin{textblock*}{21cm}(1cm,1cm)
\textbf{\sffamily INSTITUTE FOR GLOBAL CITY POLICING}
\end{textblock*}

\begin{textblock*}{18cm}(3cm, 13.66cm)
\raggedright \sffamily
\begin{singlespace}
\colorbox{white}{\hspace{1cm}\parbox[c][5.9cm]{16cm}{
{\fontsize{40}{32}\selectfont \bfseries \mbox{Stop and search}\\\mbox{in London}
\vspace{6pt}}

{\fontsize{32}{26}\selectfont \mbox{`r report_period`} }

}\hspace{1cm}}
\end{singlespace}
\end{textblock*}

\begin{textblock*}{11.43cm}(0cm, 24.13cm)
\colorbox{uclyellow}{\parbox[c][2.63cm]{\textwidth}{
\centering \bfseries \sffamily \fontsize{16}{16}\selectfont 
Dr Matt Ashby\ \ |\ \ `r format(Sys.time(), "%B %Y")`
}}
\end{textblock*}

\ 

\thispagestyle{empty}
\newpage



# Main points



\textbf{\sffamily `r headline1`}

\textbf{\sffamily `r headline2`}

\textbf{\sffamily `r headline4`}



```{r Twitter cover image, include=FALSE}
cover_bg <- str_glue("output/{report_q}/cover_image_{report_q}.jpg") %>% 
  here::here() %>% 
  image_read() %>% 
  image_scale("1200x") %>% 
  image_crop("x800", gravity = "West")
cover_logo <- here::here("output/ucl-banner-land-yellow-rgb.png") %>% 
  image_read() %>% 
  image_scale("1200x")

twitter_cover <- image_composite(cover_bg, cover_logo) %>% 
  image_annotate("INSTITUTE FOR GLOBAL CITY POLICING", location = "+17+17", 
                 font = "Arial", size = 13.5, weight = 700) %>% 
  image_draw()
rect(200, 280, 1200, 560, col = "white", border = NA)
text(235, 370, "Stop and search", adj = c(0, 0), family = "Arial", cex = 6, font = 2)
text(235, 405, "in London", adj = c(0, 0.5), family = "Arial", cex = 6, font = 2)
text(235, 440, report_period, adj = c(0, 1), family = "Arial", cex = 5)
dev.off()

image_write(
  twitter_cover, 
  here::here(str_glue("output/{report_q}/{report_q}_00_cover.png"))
)
```



# Introduction

Stop and search is a legal power that allows police officers to search people to find out if they are carrying prohibited items such as drugs, weapons or stolen goods. Stop and search means officers can confirm if a person is or is not in possession of contraband without arresting them and taking them to a police station, but it is also a source of tension between police and communities. [A review by the College of Policing](https://whatworks.college.police.uk/Research/Documents/SS_and_crime_report.pdf) found little relationship between how many searches police do and how much crime occurs, but [police insist stop and search helps them fight crime](https://www.met.police.uk/advice/advice-and-information/st-s/stop-and-search/why-we-use-stop-and-search/). _Stop and Search in London_ reports analyse stop and search in London over the past 12 months and are updated every three months.



(ref:chart-trend-overall) Number of stop-and-searches in London, `r format(min(stops$date), '%B %Y')` to `r {format(max(stops$date), '%B %Y')}`

```{r chart-trend-overall, fig.asp=0.5, fig.cap="(ref:chart-trend-overall)", fig.pos="bh", message=FALSE, warning=FALSE}
chart_trend <- monthly_counts_anomaly %>% 
  filter(year_month >= ymd("2018-08-01")) %>% 
  ggplot(aes(year_month, observed)) + 
  # Anomalous months (part 1)
  geom_point(
    size = 3, 
    shape = 21, 
    fill = "white",
    show.legend = FALSE,
    data = filter(monthly_counts_anomaly, anomaly == "Yes")
  ) +
  # Annual counts from before 2018
  geom_rect(
    aes(
      xmin = financial_year, 
      xmax = financial_year + years(1), 
      ymin = 0, 
      ymax = n / 12
    ), 
    data = filter(historical_stops, financial_year != max(financial_year)),
    inherit.aes = FALSE,
    fill = "#EA7600",
    alpha = 0.5
  ) +
  # The final annual bar actually covers slightly more than a year because the
  # monthly data starts part-way through a financial year
  geom_rect(
    aes(
      xmin = financial_year,
      xmax = financial_year + months(16),
      ymin = 0,
      ymax = n / 12
    ),
    data = filter(historical_stops, financial_year == max(financial_year)),
    inherit.aes = FALSE,
    fill = "#EA7600",
    alpha = 0.5
  ) +
  # Monthly counts from 2018 onwards
  geom_col(fill = "#EA7600", width = 31) +
  # Anomalous months (part 2)
  ggrepel::geom_label_repel(
    aes(label = format(year_month, "%b %Y")),
    data = filter(monthly_counts_anomaly, anomaly == "Yes"),
    box.padding = unit(1, "lines"),
    direction = "x",
    xlim = c(NA, max(monthly_counts_anomaly$year_month) - years(2)),
    # segment.colour = "#EA7600",
    size = chart_elements$label_text_size,
    label.size = NA
  ) +
  geom_point(
    size = 3, 
    shape = 21, 
    fill = NA,
    show.legend = FALSE,
    data = filter(monthly_counts_anomaly, anomaly == "Yes")
  ) +
  annotate(
    "text", 
    x = min(historical_stops$financial_year) + months(3),
    y = (max(historical_stops$n) / 12) * 0.1,
    label = "until 2018, only annual counts of\nsearches were published by the Home Office",
    colour = "grey20",
    size = chart_elements$label_text_size,
    lineheight = chart_elements$label_text_lineheight,
    hjust = 0,
    vjust = 1
  ) +
  # Labels for this year
  geom_label(
    data = tibble(
      year_month = mid_point(date_start, date_end),
      observed = 0,
      legislation = "searches under PACE s. 1\n(based on reasonable suspicion)"
    ),
    colour = chart_elements$label_text_colour,
    fill = rgb(1, 1, 1, 0.6),
    hjust = 0.5,
    label = "past\n12\nmonths",
    label.padding = unit(0.1, "lines"),
    label.size = NA,
    lineheight = 0.9,
    size = chart_elements$label_text_size * 0.9,
    vjust = 0
  ) +
  annotate(
    "segment", 
    x = date_start, 
    xend = date_end, 
    y = 0, 
    yend = 0, 
    colour = chart_elements$label_line_colour
  ) +
  annotate(
    "point", 
    x = date_start, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  annotate(
    "point", 
    x = date_end, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  # End of labels for this year
  coord_cartesian(clip = "off") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(labels = comma_format(), limits = c(0, NA), 
                     expand = expansion(mult = c(0, 0.025)), 
                     position = "right") +
  scale_colour_manual(values = c("No" = "grey70", "Yes" = "#EA7600"), 
                      guide = guide_none()) +
  scale_alpha_manual(
    values = c(`TRUE` = 1),
    labels = c(`TRUE` = "trend, excluding anomalies")
  ) +
  labs(
    x = NULL, 
    y = "number of searches per month", 
    alpha = NULL
  ) +
  theme_stop_search() +
  theme(
    axis.title.y = element_text(hjust = 0),
    legend.background = element_rect(colour = NA, fill = "white"),
    legend.justification = c(1, 1),
    legend.position = c(1, 1)
  )

save_chart(
  chart_trend, 
  "01_trend",
  "Number of stop-and-searches in London over time"
)

chart_trend
```



**Between `r report_between`, police officers in London carried out `r comma(pluck(count(filter(stops, date >= date_start)), "n", 1))` stop-and-searches**, or about `r comma(pluck(count(filter(stops, date >= date_start)), "n", 1) / 52)` per week. Of those, `r percent(prop_of_total(stops, force, force == "Metropolitan Police Service"))` were conducted by the Metropolitan Police, `r percent(prop_of_total(stops, force, force == "British Transport Police"))` by British Transport Police and `r ifelse(prop_of_total(stops, force, force == "City of London Police") < 0.001, "less than 0.1%", percent(prop_of_total(stops, force, force == "City of London Police")))` by City of London Police. Across the three forces, `r percent(prop_of_total(stops, type, type == "Person search"))` of stops were of pedestrians, `r percent(prop_of_total(stops, type, type == "Person and Vehicle search"))` of people in vehicles and `r percent(prop_of_total(stops, type, type == "Vehicle search"))` of only vehicles.



```{r}
rolling_counts <- mutate(
  monthly_counts, 
  annual = slider::slide_int(n, sum, .before = 11, .complete = TRUE),
  annual_diff = annual - lag(annual, n = 12)
)

annual_change <- with(
  rolling_counts,
  str_glue(
    ifelse(nth(annual, -1) - nth(annual, -13) > 0, "increase", "decrease"),
    " of ",
    percent(abs((nth(annual, -1) - nth(annual, -13)) / nth(annual, -13)))
  )
)

previous_annual_change <- with(
  rolling_counts,
  str_glue(
    ifelse(nth(annual, -13) - nth(annual, -24) > 0, "increase", "decrease"),
    " of ",
    percent(abs((nth(annual, -13) - nth(annual, -24)) / nth(annual, -24)))
  )
)
```



The number of searches carried out in `r report_period` was **a year-on-year `r annual_change`** (Figure \@ref(fig:chart-trend-overall)), compared to an annual `r previous_annual_change` in the 12 months prior to that. Prior to 2018, the number of searches had decreased every year since 2009, dropping by `r percent(abs((last(historical_stops$n) - first(historical_stops$n)) / first(historical_stops$n)))` in nine years.



# What items are people searched for?

(ref:chart-search-types) Searches by type of object being searched for, `r report_period`

```{r chart-search-types, fig.asp=0.33, fig.cap="(ref:chart-search-types)"}
chart_search_types <- searches_by_type %>% 
  filter(as_date(year_month) >= date_start) %>% 
  count(object, wt = n) %>% 
  mutate(
    align = ifelse(n > max(n) / 2, 1, 0),
    object = fct_relevel(fct_reorder(object, n), "unknown", "other"),
    percent = percent(n / sum(n), accuracy = 0.1)
  ) %>% 
  ggplot(aes(n, object, fill = object, label = str_glue(" {percent} "), 
             hjust = align)) +
  geom_col() +
  geom_text(size = chart_elements$label_text_size) +
  scale_x_continuous(
    labels = comma_format(), 
    expand = expansion(mult = c(0, 0.025))
  ) +
  scale_fill_manual(values = search_type_colours) +
  labs(
    x = "number of searches", 
    y = NULL
  ) +
  theme_stop_search() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank()
  )

save_chart(
  chart_search_types, 
  "02_types",
  str_glue("Searches by type of object being searched for, {report_period}"),
)

chart_search_types
```



```{r search types change}
search_type_trends <- stops %>% 
  count(year_month, object) %>% 
  filter(
    !object %in% c("other", "unknown"), 
    as_date(year_month) >= date_start
  ) %>% 
  arrange(object, year_month) %>% 
  mutate(change = n / first(n)) %>% 
  as_tsibble(key = object, index = year_month) %>% 
  model(lm = TSLM(change ~ trend())) %>% 
  mutate(coefs = map(lm, broom::tidy)) %>% 
  unnest(cols = coefs) %>% 
  filter(term == "trend()") %>% 
  mutate(abs_term = abs(estimate)) %>% 
  arrange(desc(abs_term))

search_type_increase <- search_type_trends %>% 
  filter(estimate > 0 & p.value < 0.05) %>% 
  pull("object")

search_type_decrease <- search_type_trends %>% 
  filter(estimate < 0 & p.value < 0.05) %>% 
  pull("object")

search_type_changes <- case_when(
  length(search_type_increase) == length(main_types) ~
    paste(to_text(search_type_increase), "have all increased"),
  length(search_type_decrease) == length(main_types) ~
    paste(to_text(search_type_decrease), "have all decreased"),
  length(search_type_increase) > 0 & length(search_type_decrease) > 0 ~
    paste(
      to_text(search_type_increase), "have increased while searches for",
      to_text(search_type_decrease), "have decreased"
    ),
  length(search_type_increase) > 0 ~
    paste(to_text(search_type_increase), "have significantly increased"),
  length(search_type_decrease) > 0 ~ 
    paste(to_text(search_type_decrease), "have significantly decreased"), 
  TRUE ~ NA_character_
)

search_type_changes <- ifelse(
  !is.na(search_type_changes),
  paste(
    "In the past 12 months, the number of searches for",
    search_type_changes,
    "(Figure \\@ref(fig:chart-search-types-change))."
  ),
  ""
)

most_common_search_types <- searches_by_type %>% 
  filter(as_date(year_month) >= date_start) %>% 
  count(object, wt = n) %>% 
  janitor::adorn_percentages(denominator = "col")

prop_weapons_text <- prop_weapons_model %>% 
  filter(term == "trend()") %>% 
  # If the proportion of searches that were for weapons changed by an annualised
  # 5% or more *and* that change was significant, describe that change
  mutate(text = case_when(
    estimate * 12 > 0.05 & p.value < 0.05 ~ 
      " Over that time, the proportion of searches that were for weapons increased.",
    estimate * 12 < -0.05 & p.value < 0.05 ~ 
      " Over that time, the proportion of searches that were for weapons decreased.",
    TRUE ~ ""
  )) %>% 
  pluck("text")
```



Police officers are empowered to search people for different items -- including drugs, items to use in theft or criminal damage, stolen goods, weapons and even some fireworks -- under different acts of parliament. Although police emphasise that stop and search "[protects Londoners by taking weapons off the streets](https://www.met.police.uk/police-forces/metropolitan-police/areas/about-us/about-the-met/stop-and-search/)", only about one in `r number_to_text(round(1 / pluck(filter(most_common_search_types, object == "weapons"), "n")))` searches between `r report_between` were for weapons -- **`r percent(pluck(slice_max(most_common_search_types, n), "n"))` of searches were for `r pluck(slice_max(most_common_search_types, n), "object")`** (Figure \@ref(fig:chart-search-types)).`r prop_weapons_text`



(ref:chart-search-types-change) Change in number of searches by type, `r format(min(stops$date), '%B %Y')` to `r format(max(stops$date), '%B %Y')`

```{r chart-search-types-change, fig.asp=0.5, fig.cap="(ref:chart-search-types-change)", fig.pos="bh"}
search_type_order <- searches_by_type %>% 
  filter(object %in% main_types) %>% 
  group_by(object) %>% 
  summarise(n = mean(n), .groups = "drop") %>% 
  arrange(desc(n)) %>% 
  pull(object)

chart_types_change <- stops %>%
  filter(
    object %in% main_types, 
    year_month >= yearmonth(ymd("2018-08-01"))
  ) %>%
  count(year_month, object) %>% 
  group_by(object) %>% 
  arrange(year_month) %>% 
  mutate(
    change = n / first(n),
    year_month = as_date(year_month)
  ) %>% 
  ungroup() %>% 
  mutate(object = fct_relevel(object, search_type_order)) %>% 
  ggplot(aes(year_month, n, colour = object)) +
  geom_segment(aes(xend = year_month, yend = 0), colour = "grey70", 
               size = 0.25) +
  geom_point(colour = "grey70", size = 0.75) +
  geom_smooth(method = "loess", formula = "y ~ x", se = FALSE) +
  # Labels for this year
  geom_label(
    data = tibble(
      year_month = mid_point(date_start, date_end),
      n = 0,
      object = suppressWarnings(fct_relevel(last(search_type_order), search_type_order))
    ),
    colour = chart_elements$label_text_colour,
    fill = rgb(1, 1, 1, 0.6),
    hjust = 0.5,
    label = "past\n12\nmonths",
    label.padding = unit(0.1, "lines"),
    label.size = NA,
    lineheight = 0.9,
    size = chart_elements$label_text_size * 0.9,
    vjust = 0
  ) +
  annotate(
    "segment", 
    x = date_start, 
    xend = date_end, 
    y = 0, 
    yend = 0, 
    colour = chart_elements$label_line_colour
  ) +
  annotate(
    "point", 
    x = date_start, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  annotate(
    "point", 
    x = date_end, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  # End of labels for this year
  coord_cartesian(clip = "off") +
  scale_x_date(date_breaks = "1 year", date_labels = "'%y",
               expand = expansion(mult = c(0.025, 0.025))) +
  scale_y_continuous(
    labels = comma_format(accuracy = 1), 
    limits = c(0, NA),
    position = "right"
  ) +
  scale_colour_manual(values = search_type_colours) +
  facet_wrap(
    facets = vars(object), 
    nrow = 1,
    scales = "free", 
    labeller = as_labeller(function (string) paste("searches for\n", string))
  ) +
  labs(
    x = NULL, 
    y = "number of searches per month"
  ) +
  theme_stop_search() +
  theme(
    axis.title.y = element_text(hjust = 0),
    legend.position = "none"
  )

save_chart(
  chart_types_change,
  "03_types_change",
  str_glue(
    "Change in number of searches by type, ",
    "{format(min(stops$date), '%B %Y')} to {format(max(stops$date), '%B %Y')}"
  )
)

chart_types_change
```



About `r percent(sum(pull(filter(most_common_search_types, object %in% main_types), "n")))` of searches are looking for the four main types of contraband: `r vector_to_text(main_types)`. `r search_type_changes`



```{r weapons trend, include=FALSE}
weapon_search_counts <- stops %>% 
  filter(
    legislation %in% c(
      "Police and Criminal Evidence Act 1984 (section 1)", 
      "Criminal Justice and Public Order Act 1994 (section 60)"
    ),
    object_of_search %in% 
      c("offensive weapons", "anything to threaten or harm anyone"),
    year_month >= yearmonth(ymd("2018-06-01"))
  ) %>% 
  count(legislation, year_month) %>% 
  mutate(
    year_month = as_date(year_month),
    diff = n - lag(n),
    diff_perc = diff / lag(n)
  )

weapon_search_annual_counts <- weapon_search_counts %>% 
  filter(year_month >= date_start) %>% 
  count(legislation, wt = n) %>% 
  mutate(prop = n / sum(n))
  
weapon_search_anomolies <- weapon_search_counts %>% 
  group_by(legislation) %>% 
  nest() %>% 
  mutate(
    anomalies = map(data, timetk::tk_anomaly_diagnostics, 
                    .date_var = year_month, .value = n)
  ) %>% 
  unnest(cols = anomalies)

weapon_search_trends <- weapon_search_counts %>% 
  filter(year_month >= max(year_month) - months(11)) %>% 
  group_by(legislation) %>% 
  arrange(year_month) %>% 
  mutate(
    change = n / first(n),
    year_month = as_date(year_month)
  ) %>% 
  ungroup() %>% 
  as_tsibble(key = legislation, index = year_month) %>% 
  model(lm = fable::TSLM(change ~ trend())) %>% 
  mutate(coefs = map(lm, broom::tidy)) %>% 
  unnest(cols = c(coefs)) %>% 
  filter(term == "trend()")
```



Police can search people for weapons using two different legal powers. Searches under [section 1 of the Police and Criminal Evidence Act 1984](https://www.legislation.gov.uk/ukpga/1984/60/section/1) (PACE) require the officer to have "reasonable grounds for suspecting" that the person is carrying an offensive weapon or other prohibited item. Conversely, officers can search people under [section 60 of the Criminal Justice and Public Order Act 1994](https://www.legislation.gov.uk/ukpga/1994/33/section/60) (CJPOA) without having any reason to think the person has a weapon, as long as a more-senior officer believes "incidents involving serious violence may take place" in the area. These 'section 60' searches are particularly controversial because they allow officers to search *anyone* in an area, even if there is no reason to think they have a weapon in their possession. Between `r report_between`, `r percent(pull(filter(weapon_search_annual_counts, legislation == "Police and Criminal Evidence Act 1984 (section 1)"), prop))` of weapons searches were based on reasonable suspicion under PACE section 1, with the remaining `r percent(pull(filter(weapon_search_annual_counts, legislation == "Criminal Justice and Public Order Act 1994 (section 60)"), prop))` (`r comma(pull(filter(weapon_search_annual_counts, legislation == "Criminal Justice and Public Order Act 1994 (section 60)"), n))` searches) were conducted without the need for suspicion based on authorisations under CJPOA section 60. Police do not publish any information about authorisations made under section 60 so it is difficult to track any patterns or trends, although section-60 searches are typically higher in August due to the Notting Hill Carnival, which was cancelled in 2020.



(ref:chart-trend-weapons) Change in number of searches for weapons, `r format(min(stops$date), '%B %Y')` to `r format(max(stops$date), '%B %Y')`

```{r chart-trend-weapons, fig.asp=0.5, fig.cap="(ref:chart-trend-weapons)"}
weapon_search_chart_data <- weapon_search_anomolies %>% 
  # Remove negative anomalies
  mutate(anomaly = ifelse(remainder < 0, "No", anomaly)) %>% 
  group_by(legislation, anomaly) %>% 
  mutate(
    anomaly_rank = min_rank(desc(abs(remainder) / observed)),
    legislation = recode_factor(
      legislation,
      "Police and Criminal Evidence Act 1984 (section 1)" = 
        "searches under PACE s. 1\n(based on reasonable suspicion)", 
      "Criminal Justice and Public Order Act 1994 (section 60)" = 
        "searches under CJPOA s. 60\n(based on authorisation)"
    )
  )

chart_trend_weapons <- ggplot(weapon_search_chart_data, 
       aes(year_month, observed, colour = legislation)) +
  geom_segment(aes(xend = year_month, yend = 1), size = 0.25, 
               colour = "grey70") +
  ggrepel::geom_label_repel(
    aes(label = format(year_month, "%b %Y")),
    data = filter(weapon_search_chart_data, anomaly == "Yes", anomaly_rank <= 4),
    box.padding = unit(1, "lines"),
    # direction = "x",
    fill = rgb(1, 1, 1, 0.5),
    force = 10,
    label.size = NA,
    size = chart_elements$label_text_size,
    ylim = c(5000, NA)
  ) +
  geom_point(
    data = filter(weapon_search_chart_data, anomaly == "Yes", anomaly_rank <= 4),
    size = 3, 
    shape = 21, 
    fill = "white", 
    show.legend = FALSE
  ) +
  geom_point(size = 0.75) +
  geom_smooth(
    method = "loess", 
    formula = "y ~ x", 
    se = FALSE, 
    data = filter(weapon_search_chart_data, anomaly == "No")
  ) +
  # Labels for this year
  geom_label(
    data = tibble(
      year_month = mid_point(date_start, date_end),
      observed = 0,
      legislation = "searches under PACE s. 1\n(based on reasonable suspicion)"
    ),
    colour = chart_elements$label_text_colour,
    fill = rgb(1, 1, 1, 0.6),
    hjust = 0.5,
    label = "past 12 months",
    label.padding = unit(0.1, "lines"),
    label.size = NA,
    lineheight = 0.9,
    size = chart_elements$label_text_size * 0.9,
    vjust = 0
  ) +
  annotate(
    "segment", 
    x = date_start, 
    xend = date_end, 
    y = 0, 
    yend = 0, 
    colour = chart_elements$label_line_colour
  ) +
  annotate(
    "point", 
    x = date_start, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  annotate(
    "point", 
    x = date_end, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  # End of labels for this year
  coord_cartesian(clip = "off") +
  scale_x_date(
    date_breaks = "6 months", 
    date_labels = "%b '%y", 
    expand = expansion(mult = c(0.025, 0.025))
  ) +
  scale_y_continuous(
    labels = comma_format(), 
    limits = c(0, NA), 
    expand = expansion(mult = c(0, 0.05)), 
    position = "right"
  ) +
  scale_colour_manual(
    values = c(ucl_colours_list[["Light Red"]], 
               ucl_colours_list[["Bright Pink"]])
  ) +
  facet_grid(
    cols = vars(legislation), 
    scales = "free"
  ) +
  labs(
    caption = "highlighted months are anomalies with unusually high numbers of stops",
    x = NULL, 
    y = "number of searches"
  ) +
  theme_stop_search() +
  theme(
    axis.title.y = element_text(hjust = 0),
    legend.position = "none"
  )

save_chart(
  chart_trend_weapons, 
  "04_trend_weapons",
  str_glue(
    "Change in number of searches for weapons, ",
    "{format(min(stops$date), '%B %Y')} to {format(max(stops$date), '%B %Y')}"
  )
)

chart_trend_weapons
```



```{r}
# SECTION 1
weapon_trend_s1 <- weapon_search_trends %>% 
  filter(legislation == "Police and Criminal Evidence Act 1984 (section 1)") %>% 
  mutate(text = ifelse(
    p.value < 0.05,
    str_glue(
      ifelse(estimate > 0, "increased", "decreased"), " by about ",
      percent(abs(median(pull(filter(weapon_search_counts, legislation == "Police and Criminal Evidence Act 1984 (section 1)", year_month >= max(year_month) - months(11)), diff_perc)))),
      " per month on average"
    ),
    "not shown a significant increasing or decreasing trend"
  )) %>% pull(text)

# weapon_anomalies_s1 <- weapon_search_anomolies %>%
#   filter(
#     legislation == "Police and Criminal Evidence Act 1984 (section 1)",
#     anomaly == "Yes",
#     yearquarter(year_month) == max(yearquarter(year_month))
#   ) %>% 
#   mutate(
#     anomaly_text = ifelse(
#       remainder > 0, 
#       str_glue("anomalously high in {month(year_month, label = TRUE, abbr = FALSE)}"), 
#       str_glue("anomalously low in {month(year_month, label = TRUE, abbr = FALSE)}")
#     )
#   ) %>% 
#   pull(anomaly_text) %>% 
#   {
#     case_when(
#       length(.) == 3 ~ vector_to_text(.),
#       length(.) > 0 ~ as.character(str_glue("{vector_to_text(.)} and within the expected range otherwise")),
#       TRUE ~ "within the expected range"
#     )
#   }

weapon_previous_anomalies_s1 <- weapon_search_anomolies %>% 
  filter(
    anomaly == "Yes", 
    legislation == "Police and Criminal Evidence Act 1984 (section 1)"
  ) %>% 
  mutate(high = remainder > 0) %>% 
  {
    
    anm <- case_when(
      sum(.$high) > 0 & sum(!.$high) > 0 ~ str_glue(
        "anomalously high in ",
        "{vector_to_text(format(.$year_month[.$high], '%B %Y'))} ",
        "but anomalously low in ",
        "{vector_to_text(format(.$year_month[!.$high], '%B %Y'))}"
      ),
      sum(.$high) > 0 ~ str_glue(
        "anomalously high in ",
        "{vector_to_text(format(.$year_month[.$high], '%B %Y'))}"
      ),
      sum(!.$high) > 0 ~ str_glue(
        "anomalously low in ",
        "{vector_to_text(format(.$year_month[!.$high], '%B %Y'))}"
      ),
      TRUE ~ ""
    )
    
    ifelse(
      anm != "", 
      str_glue("In comparison to that trend, the number of these searches was {anm}"),
      ""
    )
    
  }


# SECTION 60
weapon_trend_s60 <- weapon_search_trends %>% 
  filter(legislation == "Criminal Justice and Public Order Act 1994 (section 60)") %>% 
  mutate(text = ifelse(
    p.value < 0.05,
    str_glue(
      ifelse(estimate > 0, "increased", "decreased"), " by about ",
      percent(abs(median(pull(filter(weapon_search_counts, legislation == "Criminal Justice and Public Order Act 1994 (section 60)", year_month >= max(year_month) - months(11)), diff_perc)))),
      " per month on average"
    ),
    ifelse(
      weapon_trend_s1 == "not shown a significant increasing or decreasing trend",
      "also not shown a significant increasing or decreasing trend",
      "not shown a significant increasing or decreasing trend"
    )
  )) %>% pull(text)

# weapon_anomalies_s60 <- weapon_search_anomolies %>%
#   filter(
#     legislation == "Criminal Justice and Public Order Act 1994 (section 60)",
#     anomaly == "Yes",
#     yearquarter(year_month) == max(yearquarter(year_month))
#   ) %>% 
#   mutate(
#     anomaly_text = ifelse(
#       remainder > 0, 
#       str_glue("anomalously high in {month(year_month, label = TRUE, abbr = FALSE)}"), 
#       str_glue("anomalously low in {month(year_month, label = TRUE, abbr = FALSE)}")
#     )
#   ) %>% 
#   pull(anomaly_text) %>% 
#   {
#     case_when(
#       length(.) == 3 ~ as.character(str_glue("{vector_to_text(.)} compared to that trend")),
#       length(.) > 0 ~ as.character(str_glue("{vector_to_text(.)} compared to that trend, and within the expected range otherwise")),
#       TRUE ~ "within the range that would be expected based on that trend"
#     )
#   }

weapon_previous_anomalies_s60 <- weapon_search_anomolies %>% 
  filter(
    anomaly == "Yes", 
    legislation == "Criminal Justice and Public Order Act 1994 (section 60)"
  ) %>% 
  mutate(high = remainder > 0) %>% 
  {
    
    anm <- case_when(
      sum(.$high) > 0 & sum(!.$high) > 0 ~ str_glue(
        "anomalously high in ",
        "{vector_to_text(format(.$year_month[.$high], '%B %Y'))} ",
        "but anomalously low in ",
        "{vector_to_text(format(.$year_month[!.$high], '%B %Y'))}"
      ),
      sum(.$high) > 0 ~ str_glue(
        "anomalously high in ",
        "{vector_to_text(format(.$year_month[.$high], '%B %Y'))}"
      ),
      sum(!.$high) > 0 ~ str_glue(
        "anomalously low in ",
        "{vector_to_text(format(.$year_month[!.$high], '%B %Y'))}"
      ),
      TRUE ~ ""
    )
    
    ifelse(
      anm != "", 
      str_glue(", with searches having been {anm}"),
      ""
    )
    
  }
```



Searches based on reasonable suspicion the person being searched is carrying a weapon have `r weapon_trend_s1` over the past 12 months (Figure \@ref(fig:chart-trend-weapons)). `r weapon_previous_anomalies_s1`.
No-suspicion searches under section 60 have `r weapon_trend_s60` over the past 12 months`r weapon_previous_anomalies_s60`.



# Who do police search?

Of the `r comma(pull(count(filter(person_stops, date >= date_start)), n))` searches of pedestrians and vehicle occupants from `r report_period`, **`r percent(pull(filter(tabyl(filter(person_stops, gender %in% c("female", "male"), date >= date_start), gender), gender == "male"), percent))` were searches of men or boys**. Of all people searched, `r percent(sum(pull(filter(age_counts, age_range %in% c("under 10", "10-17")), percent)))` were aged under 18, `r percent(pull(filter(age_counts, age_range == "18-24"), percent))` were between 18 and 24, and `r percent(sum(pull(filter(age_counts, age_range %in% c("25-34", "over 34")), percent)))` were 25 or older. The self-defined ethnicity of the person searched was known for `r percent(1 - pull(filter(sde_counts, is.na(sde_group)), percent))` of searches, of which `r percent(pull(slice(sde_counts, 1), valid_percent))` of people described themselves as `r pull(slice(sde_counts, 1), sde_group)`, `r percent(pull(slice(sde_counts, 2), valid_percent))` as `r pull(slice(sde_counts, 2), sde_group)` and `r percent(pull(slice(sde_counts, 3), valid_percent))` as `r pull(slice(sde_counts, 3), sde_group)`.

**Search rates vary hugely across different groups**. Of the `r nrow(stops_disparity)` combinations of sex, age and self-defined ethnicity present in the search data, `r nrow(filter(stops_disparity, search_rate > overall_search_rate))` groups were searched at a higher rate than the rate for the population as a whole (Figure \@ref(fig:chart-disparity)). While disparity between ethnic groups has generated much comment, `r ifelse(length(stops_model) > 0, str_glue("{vector_to_text(stops_model)} are more-powerful predictors of a group having a higher search rate than that group being non-white"), "there are also disparities between age groups and sexes")`. The reasons for these differences are likely to be complex: many types of offending are concentrated among some groups (particularly young men) as well as in some neighbourhoods, and there are [longstanding issues of bias and stereotyping among police and in society](https://www.bbc.co.uk/news/uk-47300343). There is also an interaction between factors such as deprivation and the amount of time people spend in public (where almost-all searches occur). There is no way to know from the data analysed here what combination of these factors drives the disparities in search rates.



(ref:chart-disparity) Search rates for different demographic groups, `r report_period`

```{r chart-disparity, fig.asp=0.5, fig.cap="(ref:chart-disparity)"}
chart_disparity <- stops_disparity %>% 
  arrange(desc(search_rate)) %>% 
  filter(search_rate > overall_search_rate) %>% 
  add_row(group = "**all people**", search_rate = overall_search_rate) %>% 
  mutate(
    group = fct_relevel(fct_reorder(group, search_rate), "**all people**", 
                        after = Inf)
  ) %>% 
  ggplot(aes(search_rate, group, fill = group == "**all people**")) +
  geom_col(width = 0.75) +
  geom_vline(
    aes(xintercept = x),
    data = tibble(
      x = seq(0, ceiling(max(stops_disparity$search_rate)), by = 10)
    ),
    colour = "white",
    alpha = 0.75
  ) +
  scale_x_continuous(expand = c(0, 0.05)) +
  scale_fill_manual(
    values = c(`TRUE` = "black", `FALSE` = ucl_colours_list[["Orange"]])
  ) +
  labs(
    x = "searches per 1,000 people"
  ) +
  theme_stop_search() +
  theme(
    axis.text.y = element_markdown(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

save_chart(
  chart_disparity, 
  "05_disparity",
  str_glue("Search rates for different demographic groups, {report_period}"),
)

chart_disparity
```



```{r}
stops_disparity_object_chart <- stops_disparity_object %>%
  group_by(object) %>%
  mutate(
    rel_search_rate = search_rate / (sum(searches) / (sum(people) / 1000))
  ) %>%
  ungroup() %>%
  arrange(desc(rel_search_rate)) %>%
  mutate(
    object_label = recode(
      object,
      "weapons (based on authorisation)" = "weapons (s. 60)",
      "weapons (based on reasonable suspicion)" = "weapons (s. 1)"
    ),
    group = str_glue("{group}"),
    group = fct_reorder(group, rel_search_rate)
  )

disparity_groups_text <- stops_disparity_object_chart %>% 
  arrange(desc(rel_search_rate)) %>% 
  group_by(object) %>% 
  slice(1) %>% 
  ungroup() %>% 
  {
    
    text <- case_when(
      # The same age-sex-ethnicity group has the highest disparity for all 
      # search objects
      length(unique(.$group)) == 1 ~ 
        str_glue(
          "{.$age_label[1]} aged {.$age_range[1]} who ",
          "identified as {.$ethnic_group[1]} experienced the highest rate of ",
          "searches for all the five main types of search"
        ),
      # The same age-ethnicity group has the highest disparity for all search 
      # objects
      length(unique(.$sex)) == 1 & length(unique(.$ethnic_group)) == 1 ~ 
        str_glue(
          "{.$age_label[1]} who identified as {.$ethnic_group[1]} experienced ",
          "the highest rate of searches for all the five main types of search"
        ),
      # The same age-sex group has the highest disparity for all search objects
      length(unique(.$sex)) == 1 & length(unique(.$age_range)) == 1 ~ 
        str_glue(
          "{.$age_label[1]} aged {.$age_range[1]} experienced the highest ",
          "rate of searches for all the five main types of search"
        ),
      # An age-sex-ethnicity group has the highest disparity for two or more 
      # search objects
      pluck(count(., age_range, ethnic_group, sex, sort = TRUE), "n", 1) >= 3 ~
        str_glue(
          "{.$age_label[1]} aged {.$age_range[1]} who identified as ",
          "{.$ethnic_group[1]} experienced the highest rate of searches for ",
          "{number_to_text(pluck(arrange(count(., age_range, ethnic_group, sex), desc(n)), 'n', 1))}",
          "of the five main types of search"
        ),
      # A sex-ethnicity group has the highest disparity for two or more search 
      # objects
      pluck(count(., ethnic_group, sex, sort = TRUE), "n", 1) >= 3 ~
        str_glue(
          "{.$age_label[1]} who identified as {.$ethnic_group[1]} experienced ",
          "the highest rate of searches for ",
          "{number_to_text(pluck(arrange(count(., age_range, ethnic_group, sex), desc(n)), 'n', 1))} ",
          "of the five main types of search"
        ),
      # An age-sex group has the highest disparity for two or more search 
      # objects
      pluck(count(., age_range, sex, sort = TRUE), "n", 1) >= 3 ~
        str_glue(
          "{.$age_label[1]} aged {.$age_range[1]} experienced the highest ",
          "rate of searches for ",
          "{number_to_text(pluck(arrange(count(., age_range, ethnic_group, sex), desc(n)), 'n', 1))} ",
          "of the five main types of search"
        ),
      # All other cases
      TRUE ~ ""
    )
    
    ifelse(
      str_length(text) > 0,
      str_glue("Of the {number(length(unique(stops_disparity_object$sex)) * length(unique(stops_disparity_object$age_range)) * length(unique(stops_disparity_object$ethnic_group)))} combinations of age, ethnic-group and sex present in the data, {text}."),
      ""
    )
    
  }

```



In comparison to the population as a whole, `r str_remove_all(pluck(slice_max(stops_disparity, search_rate), "group"), "\\*\\*")` (the group with the highest search rate) are on-average `r number(pluck(slice_max(stops_disparity, search_rate), "search_rate") / overall_search_rate)` times more likely to be stopped and searched. Disparities in search rates also vary according to the type of search. Disparity is highest in searches for `r stops_disparity_object_chart$object[1]`, for which `r stops_disparity_object_chart$age_label[1]` aged `r stops_disparity_object_chart$age_range[1]` identifying as `r stops_disparity_object_chart$ethnic_group[1]` were `r number(stops_disparity_object_chart$rel_search_rate[1])` times more likely to be searched than the population at large. `r disparity_groups_text` 

```{r stops-disparity-adjusted, include=FALSE}
stops_disparity_adjusted <- stops_disparity %>% 
  mutate(
    age_range = factor(str_glue("{as.character(age_range)} yrs")),
    ethnic_group = fct_relevel(ethnic_group, "White")
  ) %>% 
  arrange(ethnic_group) %>% 
  group_by(age_range, sex) %>% 
  mutate(rel_rate = search_rate / first(search_rate)) %>% 
  ungroup() %>% 
  select(sex, age_range, ethnic_group, rel_rate) %>% 
  filter(ethnic_group != "White") %>% 
  mutate(
    sex_desc = case_when(
      sex == "male" & age_range == "10-17 yrs" ~ "boys",
      sex == "male" ~ "men",
      sex == "female" & age_range == "10-17 yrs" ~ "girls",
      sex == "female" ~ "women"
    ),
    group1 = str_glue("{sex_desc} aged {str_remove(age_range, ' yrs$')}"),
    group2 = str_glue("{group1} who identified as {ethnic_group}")
  ) %>% 
  arrange(desc(rel_rate))
```

It is also possible to calculate the disparity between searches specifically for different ethnic groups by comparing the rate of searches for people from ethnic minorities to white people of the same age and sex. Using this measure, `r stops_disparity_adjusted$group2[1]` were on-average `r number(stops_disparity_adjusted$rel_rate[[1]], accuracy = 0.1)` times more likely to be stopped than white `r stops_disparity_adjusted$sex_desc[1]` of the same age (Figure \@ref(fig:chart-disparity-adjusted)).

It is important to note that these disparity ratios only represent _average_ search rates for different groups -- they do not reflect the individual experience of everyone in each group. It is likely that a small number of people in each group are being searched repeatedly while others are searched far less often, but since police do not publish data on repeated searches it is difficult to know how this affects overall search rates.



```{r stops disparity object chart, eval=FALSE, fig.asp=0.75, include=FALSE}
stops_disparity_object_chart %>% 
  mutate(
    age_range = fct_relevel(age_range, "over 34", "25-34", "18-24", "10-17"),
    rate_label = number(rel_search_rate, accuracy = 0.1, prefix = "x"),
    rel_search_rate = ifelse(rel_search_rate == 0, 0.01, rel_search_rate),
    sex = recode(sex, female = "women and girls", male = "men and boys")
  ) %>% 
  ggplot(aes(object, age_range, fill = rel_search_rate, label = rate_label)) +
  geom_raster(na.rm = TRUE) +
  geom_text(size = chart_elements$label_text_size) +
  scale_x_discrete(labels = function (x) str_wrap(x, 16)) +
  scale_fill_gradient2(trans = "log", low = ucl_colours_list["Bright Blue"], 
                       high = ucl_colours_list["Bright Red"], 
                       midpoint = log(1), na.value = NA) +
  facet_grid(cols = vars(sex), rows = vars(ethnic_group), scales = "free_y", 
             switch = "y") +
  labs(
    title = str_glue("Search rates for demographic groups, relative to the ",
                     "rate for all people, {report_period}"),
    x = NULL, 
    y = NULL
  ) +
  theme_stop_search() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    panel.grid = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(hjust = 0.5, vjust = 0.5)
  )
```



(ref:chart-disparity-adjusted) Ethnic disparity in stop and search, `r report_period`

```{r chart-disparity-adjusted, fig.asp=0.4, fig.cap="(ref:chart-disparity-adjusted)"}
chart_ethnic_disparity <- stops_disparity_adjusted %>% 
  ggplot(
    aes(
      sex, 
      fct_rev(age_range), 
      fill = rel_rate, 
      label = scales::number(rel_rate, accuracy = 0.1, prefix = "x ")
    )
  ) +
  geom_tile() +
  geom_text() +
  scale_x_discrete(position = "top") +
  scale_fill_gradient2(
    trans = "log", 
    low = ucl_colours_list["Bright Blue"], 
    high = ucl_colours_list["Bright Red"], 
    midpoint = log(1), 
    na.value = NA
  ) +
  facet_grid(cols = vars(ethnic_group)) +
  labs(
    subtitle = str_glue(
      "This chart shows the rate of stop and search for people in each ethnic ",
      "group, relative to the rate of<br>stop and search for white people of ",
      "the same age and sex. Values <span style='color: #D50032;'>**greater ",
      "than 1**</span> indicate people in<br>that ethnic group are more ",
      "likely to be searched than white people of the same age and sex, ",
      "values <br><span style='color: #0097A9'>**less than 1**</span> ",
      "indicate they are less likely to be searched."
    ),
    x = NULL,
    y = NULL
  ) +
  theme_stop_search() +
  theme(
    axis.ticks = element_blank(),
    legend.position = "none",
    panel.grid = element_blank(),
    plot.title = ggtext::element_markdown(),
    plot.subtitle = ggtext::element_markdown(face = "plain", size = 10, lineheight = 1.2),
    strip.placement = "outside",
    strip.text = element_text(face = "bold")
  )

save_chart(
  chart_ethnic_disparity, 
  "disparity_ratios_adjusted", 
  str_glue("Ethnic disparity in stop-and-search in London, {report_period}"),
  height = 300, 
  width = 600, 
  title_as_subtitle = FALSE, 
  set_caption = FALSE
)

chart_ethnic_disparity
```



# How often do police find items during searches?

The purpose of stop and search is to "enable officers to allay or confirm suspicions about individuals without exercising their power of arrest" ([PACE Code A, paragraph 1.4](https://www.gov.uk/guidance/police-and-criminal-evidence-act-1984-pace-codes-of-practice)). As such, a search that does not find what is being searched for can be considered successful if it prevents an innocent person being arrested and a police officer being taken off the street unnecessarily. There is not necessarily an optimal proportion of searches that should result in the officer finding what they are looking for. Measuring outcomes is also difficult: officers may have legitimate grounds to search a group of people (e.g. all the occupants in a vehicle believed to contain a firearm) when only one person has contraband in their possession. Nevertheless, all searches are an "intrusion on the liberty of the person" (PACE Code A, paragraph 1.2) and high proportions of searches that do not find anything may indicate that searches are not well targeted.

The data released by the Home Office do not specify whether or not the item police were looking for was found during a search. Instead, we can measure whether a search resulted in some formal criminal-justice process such as an arrest. This is not a perfect measure of whether an item was found during a search, because a person might be arrested for some other reason (for example because there was an outstanding warrant for their arrest) or contraband might be found but police deal with it informally. Nevertheless, this is the least-worst measure of search outcomes that is currently available.

```{r positive results}
search_result_counts <- stops %>% 
  mutate(
    object = case_when(
      object == "weapons" & 
        legislation == "Police and Criminal Evidence Act 1984 (section 1)" ~ 
        "weapons (based on reasonable suspicion)",
      object == "weapons" & 
        legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
        "weapons (based on authorisation)",
      object == "weapons" ~ "unknown",
      TRUE ~ object
    ),
    positive = !outcome %in% c("no further action", "unknown"),
    year_month = as_date(year_month)
  ) %>% 
  filter(!object %in% c("other", "unknown"), outcome != "unknown") %>% 
  count(year_month, object, positive) %>% 
  group_by(year_month, object) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

search_result_month_counts <- stops %>% 
  mutate(positive = !outcome %in% c("no further action", "unknown")) %>% 
  count(year_month, positive) %>% 
  group_by(year_month) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

# Calculate overall hit rate for the past 12 months
hit_rate_last_year <- search_result_month_counts %>% 
  filter(as_date(year_month) >= date_start) %>% 
  count(positive, wt = n) %>% 
  janitor::adorn_percentages(denominator = "col") %>% 
  filter(positive) %>% 
  pluck("n")

# Arrange search objects in descending order of mean hit rate
search_result_order <- search_result_counts %>% 
  filter(year_month >= date_start) %>% 
  filter(positive) %>% 
  group_by(object) %>% 
  summarise(prop = median(prop), .groups = "drop") %>% 
  arrange(desc(prop))

# Calculate linear trend in hit rate over the past 12 months
search_result_trends <- search_result_counts %>% 
  filter(positive, as_date(year_month) >= date_start) %>% 
  as_tsibble(key = object, index = year_month) %>% 
  model(lm = fable::TSLM(prop ~ trend())) %>% 
  mutate(coefs = map(lm, broom::tidy)) %>% 
  unnest(cols = c(coefs)) %>% 
  filter(term == "trend()", p.value < 0.05)

# Explain whether there have been significant increases/decreases in hit rates
# for different search types over the past 12 months
search_result_trend_text <- case_when(
  # Some increased and some decreased
  nrow(filter(search_result_trends, estimate > 0)) > 0 
  & nrow(filter(search_result_trends, estimate < 0)) > 0 ~ 
    paste0(
      "for ", 
      vector_to_text(pull(filter(search_result_trends, estimate > 0), object)), 
      " resulting in a formal outcome increased while the proportion of ", 
      "searches for ", 
      vector_to_text(pull(filter(search_result_trends, estimate < 0), object)), 
      " resulting in a formal outcome decreased"
    ),
  # Some increased and none decreased
  nrow(filter(search_result_trends, estimate > 0)) > 0 ~ 
    paste0(
      "for ", 
      vector_to_text(pull(filter(search_result_trends, estimate > 0), object)), 
      " resulting in a formal outcome have", 
      ifelse(nrow(filter(search_result_trends, estimate > 0)) > 2, " all ", " "), 
      "increased"
    ),
  # Some decreased and none increased
  nrow(filter(search_result_trends, estimate < 0)) > 0 ~ 
    paste0(
      "for ", 
      vector_to_text(pull(filter(search_result_trends, estimate < 0), object)), 
      " resulting in a formal outcome have", 
      ifelse(nrow(filter(search_result_trends, estimate < 0)) > 2, " all ", " "), 
      "decreased"
    ),
  # No significant increases or decreases
  TRUE ~ "leading to a formal outcome has remained largely constant"
)

results_by_type <- stops %>% 
  mutate(
    object = case_when(
      object == "weapons" & 
        legislation == "Police and Criminal Evidence Act 1984 (section 1)" ~ 
        "weapons (based on reasonable suspicion)",
      object == "weapons" & 
        legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
        "weapons (based on authorisation)",
      object == "weapons" ~ "unknown",
      TRUE ~ object
    ),
    outcome = fct_rev(recode(
      outcome, 
      "community/local resolution" = "community resolution"
    ))
  ) %>% 
  filter(
    !object %in% c("other", "unknown"), 
    !outcome %in% c("no further action", "unknown", "drugs warning", "caution"), 
    date >= date_start
  ) %>% 
  count(object, outcome) %>% 
  group_by(object) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

most_common_outcome <- stops %>% 
  filter(!outcome %in% c("no further action", "unknown")) %>% 
  count(outcome, sort = TRUE) %>% 
  head(1) %>% 
  pull(outcome) %>% 
  as.character()
```



(ref:chart-results) Change in proportion of searches with a formal outcome, `r format(min(stops$date), '%B %Y')` to `r format(max(stops$date), '%B %Y')`

```{r chart-results, fig.asp=0.5, fig.cap="(ref:chart-results)"}
chart_hit_rate <- search_result_counts %>% 
  filter(positive) %>%
  mutate(object = fct_relevel(object, pull(search_result_order, object))) %>% 
  ggplot(aes(year_month, prop, colour = object)) +
  geom_segment(aes(xend = year_month, yend = 0), size = 0.25, 
               colour = "grey70") +
  geom_point(size = 0.75) +
  geom_smooth(method = "loess", formula = "y ~ x", se = FALSE) +
  # Labels for this year
  geom_label(
    data = tibble(
      year_month = mid_point(date_start, date_end),
      prop = 0,
      object = suppressWarnings(fct_relevel(
        first(pull(search_result_order, "object")), 
        pull(search_result_order, "object"))
      )
    ),
    colour = chart_elements$label_text_colour,
    fill = rgb(1, 1, 1, 0.6),
    hjust = 0.5,
    label = "past\n12\nmonths",
    label.padding = unit(0.1, "lines"),
    label.size = NA,
    lineheight = 0.9,
    size = chart_elements$label_text_size * 0.9,
    vjust = 0
  ) +
  annotate(
    "segment", 
    x = date_start, 
    xend = date_end, 
    y = 0, 
    yend = 0, 
    colour = chart_elements$label_line_colour
  ) +
  annotate(
    "point", 
    x = date_start, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  annotate(
    "point", 
    x = date_end, 
    y = 0, 
    colour = chart_elements$label_line_colour,
    shape = "|"
  ) +
  # End of labels for this year
  coord_cartesian(clip = "off") +
  scale_x_date(date_breaks = "1 year", date_labels = "'%y",
               expand = expansion(mult = c(0.025, 0.025))) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.025)), 
                     position = "right") +
  scale_colour_manual(values = search_type_colours) +
  facet_grid(
    cols = vars(object),
    labeller = as_labeller(function (string) {
      str_wrap(paste("searches for", string), 18)
    })
  ) +
  labs(
    x = NULL, 
    y = "proportion resulting in a formal outcome"
  ) +
  theme_stop_search() +
  theme(
    axis.title.y = element_text(hjust = 0),
    legend.position = "none"
  )

save_chart(
  chart_hit_rate, 
  "06_hit_rate",
  str_glue(
    "Change in proportion of searches with a formal outcome, ",
    "{format(min(stops$date), '%B %Y')} to {format(max(stops$date), '%B %Y')}"
  )
)

chart_hit_rate
```



Overall, about `r percent(hit_rate_last_year)` of searches between `r report_between` resulted in a formal criminal-justice outcome (arrest, charge by post, caution, fixed penalty, community/local resolution or drugs warning), while the remaining **`r percent(1 - hit_rate_last_year)` of searches resulted in no further action.** Over the past year, searches for `r head(search_result_order$object, 1)` have been most likely to lead to a formal outcome, while `r percent(1 - tail(search_result_order$prop, 1))` of searches for `r ifelse(tail(search_result_order$object, 1) == "weapons (based on authorisation)", "weapons under a section 60 authorisation", tail(search_result_order$object, 1))` resulted in no further action.

In the past 12 months, the proportion of searches `r search_result_trend_text` (Figure \@ref(fig:chart-results)). When a stop does result in formal action, the most common outcome is `r most_common_outcome` (used in `r percent(head(pull(janitor::adorn_percentages(count(filter(stops, !outcome %in% c("no further action", "unknown")), outcome, sort = TRUE), "col"), n), 1))` of cases with a formal outcome). However, which action police choose varies with the type of search: `r percent(pull(head(arrange(filter(results_by_type, outcome == most_common_outcome), desc(prop)), 1), prop))` of positive searches for `r pull(head(arrange(filter(results_by_type, outcome == most_common_outcome), desc(prop)), 1), object)` result in `r most_common_outcome`, compared to only `r percent(pull(tail(arrange(filter(results_by_type, outcome == most_common_outcome), desc(prop)), 1), prop))` of positive searches for `r pull(tail(arrange(filter(results_by_type, outcome == most_common_outcome), desc(prop)), 1), object)`. The outcomes of some searches suggest that the outcome does not relate to the type of contraband that police were looking for. For example, fixed penalties are not a legally available option for dealing with weapons or firearms offences, but `r percent(pull(filter(results_by_type, outcome == "fixed penalty", object == "weapons (based on reasonable suspicion)"), prop))` of formal outcomes to searches for weapons based on reasonable suspicion, `r percent(pull(filter(results_by_type, outcome == "fixed penalty", object == "weapons (based on authorisation)"), prop))` of formal outcomes to searches for weapons based on section-60 authorisations and `r percent(pull(filter(results_by_type, outcome == "fixed penalty", object == "firearms"), prop))` of formal outcomes to searches for firearms were fixed penalties. This suggests that some weapons and firearms searches result in police not finding weapons but discovering more-minor offences such as cannabis possession.



```{r results by type chart, eval=FALSE, fig.asp=0.67, include=FALSE}
ggplot(
  results_by_type, 
  aes(prop, outcome, fill = outcome, 
      label = str_glue(" {percent(prop, accuracy = 1)}"))
) +
  geom_col() +
  geom_text(
    colour = chart_elements$label_text_colour,
    size = chart_elements$label_text_size,
    hjust = 0
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.02)), 
    labels = percent_format()
  ) +
  scale_fill_manual(
    values = unname(ucl_colours_list[c("Bright Green", "Bright Blue", 
                                       "Bright Green", "Light Blue")])
  ) +
  facet_grid(rows = vars(object), labeller = label_wrap_gen(width = 12)) +
  labs(
    title = str_glue("Formal outcomes for different types of search, ",
                     "{report_span}"),
    x = "proportion of all searches resulting in a formal outcome",
    y = NULL
  ) +
  theme_stop_search() +
  theme(
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  )
```



```{r outcomes by ethnicity}
stops_by_ethnicity <- stops %>% 
  separate(self_defined_ethnicity, c("ethnic_group", "ethnicity"), " - ") %>% 
  mutate(
    ethnic_group = str_extract(ethnic_group, "^\\w+"),
    formal_disposal = outcome != "no further action",
    object = case_when(
      object == "weapons" & 
        legislation == "Police and Criminal Evidence Act 1984 (section 1)" ~ 
        "weapons (based on reasonable suspicion)",
      object == "weapons" & 
        legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
        "weapons (based on authorisation)",
      object == "weapons" ~ "unknown",
      TRUE ~ object
    )
  ) %>% 
  filter(
    !is.na(ethnic_group),
    !object %in% c("other", "unknown"), 
    outcome != "unknown",
    as_date(year_month) >= date_start
  )

# Count stops by ethnicity and calculate proportions
outcomes_by_ethnicity <- stops_by_ethnicity %>% 
  count(object, formal_disposal, ethnic_group) %>% 
  mutate_if(is.factor, as.character) %>% 
  group_by(object, ethnic_group) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup() %>% 
  filter(formal_disposal)

# Find order of ethnicities by hit rate for most-frequent search object
outcomes_by_ethnicity_order <- outcomes_by_ethnicity %>% 
  filter(object == pluck(slice_max(search_result_order, prop), "object")) %>% 
  arrange(prop) %>% 
  pull(ethnic_group)

# Model whether the hit rate for each search object varies between ethnicities
outcomes_by_ethnicity_model <- stops_by_ethnicity %>% 
  mutate(ethnic_group = fct_relevel(ethnic_group, "White")) %>% 
  nest(data = -object) %>% 
  mutate(model = map(data, function (x) {
    broom::tidy(glm(formal_disposal ~ ethnic_group, data = x))
  })) %>% 
  select(object, model) %>% 
  unnest(cols = model) %>% 
  filter(term %in% c("ethnic_groupAsian", "ethnic_groupBlack")) %>% 
  mutate(
    odds_ratio = exp(estimate),
    sig = p.value < 0.05 & !between(odds_ratio, 0.95, 1.05),
    search_type = str_glue("searches of {str_remove(term, '^ethnic_group')} ",
                           "people for {object}"),
    diff_type = ifelse(odds_ratio > 1, str_glue("higher for {search_type}"), 
                       str_glue("lower for {search_type}"))
  )

# Generate text to summarise these models
outcomes_by_ethnicity_model_text <- ifelse(
  sum(outcomes_by_ethnicity_model$sig) / length(outcomes_by_ethnicity_model$sig) > 0.5, 
  str_glue(
    "Just as some ethnic groups are more likely to be stopped than others, ",
    "the probability of a stop resulting in a formal criminal-justice outcome ",
    "also varies by ethnicity: over the past 12 months, the probability of a ",
    "formal outcome to a search is ",
    to_text(pull(
      arrange(filter(outcomes_by_ethnicity_model, sig), desc(odds_ratio)), 
      diff_type
    ))
  ), 
  str_glue(
    "While the rate of searches varies between ethnic groups, the probability ",
    "of a search resulting in a formal criminal-justice outcome is broadly ",
    "the same across ethnicities – over the past 12 months, ", 
    ifelse(
      sum(outcomes_by_ethnicity_model$sig) > 0, 
      str_glue(
        "the probability of a formal outcome to searches of Black or Asian ",
        "people is only significantly different from the probabilty of a ",
        "formal outcome to searches of White people for ",
        to_text(pull(filter(outcomes_by_ethnicity_model, sig), search_type))
      ),
      str_glue(
        "the probability of a formal outcome to searches of Black or Asian ",
        "people was not significantly different from the probabilty of a ",
        "formal outcome to searches of White people for any of the main ",
        "search types"
      )
    )
  )
)
```



`r outcomes_by_ethnicity_model_text` (Figure \@ref(fig:chart-outcomes-ethnicity)).



(ref:chart-outcomes-ethnicity) Proportion of searches resulting in a formal outcome, `r report_period`

```{r chart-outcomes-ethnicity, fig.asp=0.33, fig.cap="(ref:chart-outcomes-ethnicity)"}
chart_outcomes_ethnicity <- outcomes_by_ethnicity %>% 
  mutate(
    ethnic_group = fct_recode(
      fct_relevel(as.character(ethnic_group), outcomes_by_ethnicity_order),
      "Mixed*" = "Mixed",
      "Other*" = "Other"
    ),
    prop_label = percent(prop, accuracy = 1, prefix = " ", suffix = "% "),
    align = ifelse(prop > max(prop) / 2, 1, 0)
  ) %>% 
  ggplot(aes(
    x = prop, 
    y = ethnic_group, 
    colour = as.logical(align), 
    fill = object, 
    label = prop_label, 
    hjust = align
  )) +
  geom_col(colour = NA) +
  geom_text(size = chart_elements$label_text_size) +
  scale_x_continuous(breaks = c(0, 0.1, 0.2), labels = percent_format()) +
  scale_colour_manual(
    values = c(`FALSE` = chart_elements$label_text_colour, `TRUE` = "white")
  ) +
  scale_fill_manual(values = search_type_colours) +
  facet_grid(cols = vars(object), labeller = label_wrap_gen(width = 16)) +
  labs(
    x = NULL,
    y = NULL,
    caption = str_glue(
      "* should not be compared to other ethnic groups because the number of ",
      "people in these ethnic groups is small"
    )
  ) +
  theme_stop_search() +
  theme(
    axis.ticks.y = element_blank(),
    legend.position = "none",
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank()
  )

save_chart(
  chart_outcomes_ethnicity, 
  "07_chart_outcomes_ethnicity",
  str_glue("Proportion of searches resulting in a formal outcome, {report_period}") 
)

chart_outcomes_ethnicity
```



# Where do stops happen?

```{r deprivation, include=FALSE}
# Find proportion of total stops that occur in the 50% most-deprived LSOAs
stops_by_deprivation <- stops_by_lsoa %>% 
  count(lsoa_code, name = "searches") %>% 
  left_join(imd, by = "lsoa_code") %>% 
  mutate(imd_perc = ceiling(imd_perc * 100)) %>% 
  arrange(imd_perc) %>% 
  count(imd_perc, wt = searches, name = "searches") %>% 
  mutate(
    csum_searches = cumsum(searches),
    csum_searches = csum_searches / max(csum_searches)
  ) %>% 
  filter(imd_perc == 50) %>% 
  slice(1) %>% 
  pull(csum_searches)

# Find proportion of stops by type that occur in the 50% most-deprived LSOAs
stops_by_deprivation_object <- stops_by_lsoa %>% 
  filter(
    object %in% c("drugs", "firearms", "stolen goods", 
                  "weapons under section 1", "weapons under section 60")
  ) %>% 
  count(object, lsoa_code, name = "searches") %>% 
  left_join(imd, by = "lsoa_code") %>% 
  mutate(imd_perc = ceiling(imd_perc * 100)) %>% 
  arrange(imd_perc) %>% 
  count(object, imd_perc, wt = searches, name = "searches") %>% 
  group_by(object) %>% 
  mutate(
    csum_searches = cumsum(searches),
    csum_searches = csum_searches / max(csum_searches)
  ) %>% 
  filter(imd_perc == 50) %>% 
  slice(1) %>% 
  ungroup() %>% 
  arrange(desc(csum_searches))

# Construct description for concentration of stops by deprivation
stops_by_deprivation_text <- ifelse(
  stops_by_deprivation < 45, 
  str_glue(
    "Searches are also concentrated in deprived areas: ", 
    "{percent(stops_by_deprivation)} of searches took place in neighbourhoods ",
    "that were more deprived than average.", 
    ifelse(
      max(stops_by_deprivation_object$csum_searches) > stops_by_deprivation + 0.03, 
      str_glue(
        " In particular, ",
        "{percent(pull(slice(stops_by_deprivation_object, 1), csum_searches))}",
        " of searches for ",
        "{pull(slice(stops_by_deprivation_object, 1), object)} occurred in ",
        "the most-deprived half of neighbourhoods."
      ), 
      ""
    )
  ), 
  ""
)

# Construct description for concentration of stops by borough
stops_by_borough_text <- str_glue(
  pull(slice(searches_by_borough, 1), borough), 
  " (", 
  comma(pull(slice(searches_by_borough, 1), searches_per_month)),
  " searches per month), ",
  pull(slice(searches_by_borough, 2), borough),
  " (", 
  comma(pull(slice(searches_by_borough, 2), searches_per_month)),
  ") and ",
  pull(slice(searches_by_borough, 3), borough),
  " (",
  comma(pull(slice(searches_by_borough, 3), searches_per_month)),
  "), while the fewest took place in ",
  pull(slice(searches_by_borough, 33), borough),
  " (",
  comma(pull(slice(searches_by_borough, 33), searches_per_month)), 
  " searches), ",
  pull(slice(searches_by_borough, 32), borough),
  " (",
  comma(pull(slice(searches_by_borough, 32), searches_per_month)),
  ") and ",
  pull(slice(searches_by_borough, 31), borough),
  " (",
  comma(pull(slice(searches_by_borough, 31), searches_per_month)),
  ")"
)

# Construct description of places with fewer searches than the ward with the
# most searches
places_less_than_top_ward <- case_when(
  length(low_search_boroughs) > 5 ~ 
    str_glue("{number_to_text(length(low_search_boroughs))} entire boroughs"),
  length(low_search_boroughs) > 0 ~ 
    str_glue(
      "the entire boroughs of ",
      "{vector_to_text(pull(filter(searches_by_borough, searches < searches_by_ward$searches[1]), borough), final_sep = ' or ')}"
    ),
  TRUE ~ 
    str_glue(
      "the ",
      "{comma(nrow(filter(mutate(arrange(searches_by_ward, searches), csum_searches = cumsum(searches)), csum_searches < searches_by_ward$searches[1])))} ",
      "wards with the fewest searches combined"
    )
)
```



```{r prepare hotspots, include=FALSE}
# Convert data to an SF object for plotting
stops_sf <- stops %>% 
  filter(!is.na(latitude), !is.na(longitude), date >= date_start) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(27700)

# Calculate grids of 250m and 1km hexagonal cells
grid_1km <- london_outline %>% 
  st_transform(27700) %>% 
  create_grid_hexagonal(cell_size = 750) %>% 
  mutate(id = row_number())

# Count stops in each grid cell
grid_stops <- stops_sf %>% 
  st_join(grid_1km) %>% 
  as_tibble() %>% 
  count(id, name = "searches") %>% 
  full_join(grid_1km, by = "id") %>% 
  filter(!is.na(id)) %>% 
  mutate(searches = ifelse(searches == 0 | is.na(searches), NA, searches)) %>% 
  st_sf()

# Identify hotspot cells
grid_neighbours <- grid_1km %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  dnearneigh(0, 750 * 1.5) %>% 
  include.self()
grid_gi <- grid_stops %>% 
  replace_na(list(searches = 0)) %>% 
  pull(searches) %>% 
  localG(listw = nb2listw(grid_neighbours, style = "B"))
grid_stops_gi <- grid_stops %>% 
  add_column(gistar = grid_gi) %>% 
  mutate(
    # Convert the Z scores to p values
    pvalue = 2 * pnorm(-abs(as.numeric(grid_gi))),
    # Calculate if the p values are statistically significant, which by
    # convention is if p < 0.05, adjusting for multiple comparisons using 
    # p.adjustSP()
    significant = p.adjustSP(pvalue, grid_neighbours) < 0.05
  ) %>%
  select(-pvalue)

# Estimate density for grid cells
grid_kde <- stops_sf %>% 
  kde(band_width = 750 * 1.5, grid = grid_1km) %>% 
  st_drop_geometry() %>% 
  right_join(grid_stops_gi, by = "id") %>% 
  as_tibble() %>% 
  st_sf() %>% 
  st_intersection(london_outline)

# Identify nearest station for hotspot cells
hotspot_labels <- grid_kde %>% 
  filter(significant) %>% 
  arrange(desc(kde_value)) %>% 
  # Slice more rows than needed because some will be duplicates that are removed
  # later
  slice(1:50) %>% 
  st_centroid() %>% 
  mutate(
    coord_x = st_coordinates(.)[, 1], 
    coord_y = st_coordinates(.)[, 2],
    nearest_station_id = st_nearest_feature(., london_stations)
  ) %>% 
  left_join(
    st_set_geometry(london_stations, NULL), 
    by = c("nearest_station_id" = "id")
  ) %>% 
  as_tibble() %>% 
  distinct(name, .keep_all = TRUE) %>% 
  slice(1:12) %>% 
  mutate(label = str_glue("{ordinal(row_number())}. {name}"))
```



Stop and search is geographically concentrated in some parts of London: **half of searches between `r report_between` occurred in `r percent(stop_concentration)` of neighbourhoods**. `r stops_by_deprivation_text`

Of the 33 boroughs in London, the most searches in `r report_period` took place in `r stops_by_borough_text`. We can identify search hotspots by dividing London into a grid of equally-sized cells and mapping the density of searches in each grid cell (Figure \@ref(fig:chart-map)).

Of the `r comma(nrow(london_wards))` local-authority wards in London, the ward with the most searches between `r report_between` was `r searches_by_ward$ward[1]` ward in `r searches_by_ward$borough[1]`, in which there were more searches than in `r places_less_than_top_ward` (Table \@ref(tab:table-ward)).



```{r section 60 by borough}
s60_by_borough <- stops %>% 
  filter(
    !is.na(latitude), 
    !is.na(longitude), 
    date >= date_start,
    legislation == "Criminal Justice and Public Order Act 1994 (section 60)"
  ) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(27700) %>% 
  st_join(london_boroughs) %>% 
  as_tibble() %>% 
  count(borough, name = "searches") %>% 
  filter(!is.na(borough)) %>% 
  arrange(desc(searches)) %>% 
  mutate(
    perc = searches / sum(searches),
    csum_perc = cumsum(perc),
    less_than_half = row_number() == 1 | lag(csum_perc) < 0.5,
    label = str_glue("{percent(perc, accuracy = 1)} in {borough}")
  )

s60_by_borough_text <- paste0(
  "Of the ", 
  comma(nrow(filter(stops, date >= date_start, legislation == "Criminal Justice and Public Order Act 1994 (section 60)"))),
  " no-suspicion searches under section 60 from ", report_period, 
  ", more than half (",
  percent(tail(pull(filter(s60_by_borough, less_than_half), csum_perc), 1)), 
  ") took place in ",
  number_to_text(nrow(filter(s60_by_borough, less_than_half))), 
  " of the 33 boroughs in London", 
  case_when(
    # Half of cases are in 5 or fewer boroughs, in which case name them
    nrow(filter(s60_by_borough, less_than_half)) <= 5 ~
      str_glue(" ({to_text(pull(filter(s60_by_borough, less_than_half), label))})."),
    # More than 10% of searches are in one borough, in which case name it
    pluck(s60_by_borough, "perc", 1) > 0.1 ~
      str_glue(" (with ", pluck(s60_by_borough, "label", 1), " borough alone)."),
    # Otherwise don't name any boroughs
    TRUE ~ "."
  ),
  case_when(
    length(setdiff(london_boroughs$borough, s60_by_borough$borough)) > 10 ~ 
      str_glue(
        " Meanwhile, there were no section-60 searches in ",
        number_to_text(length(setdiff(
          london_boroughs$borough, 
          s60_by_borough$borough
        ))),
        " other boroughs."
      ),
    length(setdiff(london_boroughs$borough, s60_by_borough$borough)) > 1 ~ 
      str_glue(
        " Meanwhile, there were no section-60 searches in ",
        to_text(
          setdiff(london_boroughs$borough, s60_by_borough$borough), 
          and = " or "
        ),
        " boroughs."
      ), 
    length(setdiff(london_boroughs$borough, s60_by_borough$borough)) == 1 ~ 
      str_glue(
        " Meanwhile, there were no section-60 searches in ",
        to_text(
          setdiff(london_boroughs$borough, s60_by_borough$borough), 
          and = " or "
        )
      ), 
    TRUE ~ ""
  ),
  "."
) %>% 
  str_replace(" City of London", " the City of London")
```



Searches for weapons under section 60 can only take place in areas in which an inspector (a second-line supervisor) believes "incidents involving serious violence may take place". `r s60_by_borough_text`



```{r table-ward}
searches_by_ward %>% 
  mutate(
    searches = comma(searches, accuracy = 1),
    ward = str_glue("{row_number()}. {ward} ward, {borough}")
  ) %>% 
  select(`council ward` = ward, searches) %>% 
  head(20) %>% 
  knitr::kable(
    format = "latex",
    align = "lr", 
    booktabs = TRUE,
    caption = str_glue("Local authority wards with the ",
                       "highest number of searches, {report_period}"),
    linesep = ""
  )
```



# A note on data

This report uses data published by the Home Office at [data.police.uk](https://data.police.uk/) under the [Open Government Licence version 3.0](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/) for searches by the Metropolitan Police Service or City of London Police, or by British Transport Police in London.

Search rates are calculated using [2020 estimates of the London population by age and ethnic group](https://data.london.gov.uk/dataset/ethnic-group-population-projections) produced by the Mayor of London. Rates based on residential populations are imperfect because some people being searched in London will live outside London, but the vast majority of people searched in London are likely to also live in the region. All ethnicity figures in this report are self-defined ethnicities.

This report is published under a [Creative Commons Attribution Licence version 4.0](https://creativecommons.org/licenses/by/4.0/), meaning you are free to copy or redistribute this material in any medium or format, and to remix, transform, and build upon this material for any purpose, even commercially, as long as you comply with the licence terms.

Cover photo by [Jaanus Jagomägi on Unsplash.com](https://unsplash.com/photos/Dymu1WiZVko)



\newgeometry{top = 2cm, inner = 1cm, bottom = 3cm, outer = 1cm}

(ref:chart-map) Hotspots of searches, `r report_period`

```{r chart-map, fig.asp=1, fig.cap="(ref:chart-map)", out.width="19cm"}
chart_map <- grid_kde %>% 
  ggplot(aes(colour = kde_value, fill = kde_value)) +
  geom_sf() +
  geom_sf(data = london_boroughs, colour = "grey80", fill = NA, size = 0.2) +
  geom_sf_text(
    aes(label = str_wrap(str_replace(borough, " and ", " & "), 10)),
    data = london_boroughs,
    colour = chart_elements$label_text_colour,
    size = chart_elements$label_text_size * 0.8,
    lineheight = chart_elements$label_text_lineheight,
    inherit.aes = FALSE,
    check_overlap = TRUE
  ) +
  # West London labels to the left
  ggrepel::geom_label_repel(
    aes(x = coord_x, y = coord_y, label = str_wrap(label, 25)),
    data = filter(hotspot_labels, coord_x < median(coord_x)),
    colour = chart_elements$label_text_colour,
    fill = "white",
    size = chart_elements$label_text_size * 0.8,
    lineheight = chart_elements$label_text_lineheight,
    hjust = 1,
    box.padding = unit(2, "lines"),
    label.padding = unit(0.2, "lines"),
    label.size = NA,
    xlim = c(NA, min(hotspot_labels$coord_x) - 2500),
    max.iter = 10^5,
    max.overlaps = 100
  ) +
  # East London labels to the right
  ggrepel::geom_label_repel(
    aes(x = coord_x, y = coord_y, label = str_wrap(label, 25)),
    data = filter(hotspot_labels, coord_x >= median(coord_x)),
    colour = chart_elements$label_text_colour,
    fill = "white",
    size = chart_elements$label_text_size * 0.8,
    lineheight = chart_elements$label_text_lineheight,
    hjust = 0,
    box.padding = unit(2, "lines"),
    label.padding = unit(0.2, "lines"),
    label.size = NA,
    xlim = c(max(hotspot_labels$coord_x) + 2500, NA),
    max.iter = 10^5,
    max.overlaps = 100
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_colour_distiller(
    aesthetics = c("colour", "fill"), 
    palette = "Oranges", 
    direction = 1, 
    na.value = "grey90"
  ) +
  coord_sf(crs = 27700) +
  labs(
    colour = str_glue("number of\nsearches"),
    fill = str_glue("number of\nsearches")
  ) +
  theme_stop_search() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "none",
    panel.grid = element_blank()
  )

save_chart(
  chart_map, 
  "08_chart_map", 
  str_glue("Location of searches, {report_period}"),
  height = 600, 
  width = 600
)

chart_map

```



```{r write tweets}
tweets <- vector()

tweets[1] <- str_glue(
  "Our updated @IGCP_UCL analysis of #StopAndSearch in London covers the year ",
  "to {format(date_end, '%B %Y')} and shows the number of searches ",
  "{total_searches_change} {comma(total_searches, accuracy = 100)}.\n\n",
  "#StopSearch Report: ",
  "http://lesscrime.info/publication/stop-and-search-london-2021-q3/\n\n",
  "Thread …\n\n[{report_q}_00_cover.png]\n\nALT: Cover image for stop and ",
  "search report"
)

tweets[2] <- str_glue(
  "From {report_span}, police in London searched {comma(total_searches)}",
  " people and vehicles, or about ",
  comma(total_searches / 365, accuracy = 10),
  " each day. The number of searches has generally ",
  ifelse(monthly_counts_feats$linearity > 0, "increased", "decreased"),
  " over the past year, after substantial declines between 2010 and 2018.",
  "\n\n[{report_q}_01_trend.png]\n\nALT: Bar chart showing change in the ",
  "frequency of stop and search over time"
)

search_count_by_type <- stops %>% 
  filter(date >= date_start) %>% 
  count(object) %>% 
  janitor::adorn_percentages(denominator = "col") %>% 
  arrange(desc(n))

tweets[3] <- str_glue(
  "About {percent(pluck(search_count_by_type, 'n', 1))} of all searches are ",
  "for {pluck(search_count_by_type, 'object', 1)}, ",
  "{percent(pluck(search_count_by_type, 'n', 2))} for ",
  "{pluck(search_count_by_type, 'object', 2)} and ",
  "{percent(pluck(search_count_by_type, 'n', 3))} for ",
  "{pluck(search_count_by_type, 'object', 3)}. {search_type_changes}",
  "{report_q}_03_types_change.png]\n\nALT: Bar chart showing proportion of ",
  "searches by type\nALT: Line chart showing changes in the frequency of ",
  "different types of stop and search over time"
)

weapons_searches_by_power <- weapon_search_counts %>% 
  filter(year_month >= date_start) %>% 
  count(legislation, wt = n) %>% 
  mutate(legislation = str_extract(legislation, "\\(.+?\\)")) %>% 
  deframe()

tweets[4] <- str_glue(
  "From {report_period}, police searched ",
  "{comma(weapons_searches_by_power['(section 1)'])} people they had reason ",
  "to suspect were carrying weapons, along with ",
  "{comma(weapons_searches_by_power['(section 60)'])} people without ",
  "reasonable suspicion in areas where a mid-ranking officer believed some ",
  "people might carry weapons.\n\n[{report_q}_04_trend_weapons.png]\n\nALT: ",
  "Line chart showing changes in the frequency of searches for weapons over ",
  "time"
)

tweets[5] <- str_glue(
  "Different groups are searched at different rates – ",
  str_remove_all(
    pluck(arrange(stops_disparity, desc(search_rate)), "group", 1), 
    "\\*\\*"
  ),
  " were on-average ",
  number(pluck(arrange(stops_disparity, desc(search_rate)), "search_rate", 1) / overall_search_rate),
  " times more likely to be searched than the population at large. The ",
  "reasons for this are likely to be a complex mix of different factors, with ",
  "no single cause.\n\n[{report_q}_05_disparity.png]\n\nALT: Bar chart ",
  "showing rates at which different demographic groups are searched"
)

tweets[6] <- str_glue(
  percent(pluck(filter(search_result_annual_counts, !positive), "prop")), 
  " of searches in {report_period} resulted in no further action. In the past ",
  "two years, the proportion of searches ",
  str_replace(
    search_result_trend_text, 
    "resulting in a formal outcome", 
    "resulting in a formal outcome (e.g. arrest)"
  ),
  ".\n\n[{report_q}_06_hit_rate.png]\n\nALT: Line chart showing how the ",
  "proportion of searches resulting in a formal criminal-justice outcome has ",
  "changed over time"
)

tweets[7] <- ifelse(
  sum(outcomes_by_ethnicity_model$sig) / length(outcomes_by_ethnicity_model$sig) > 0.5, 
  str_glue(
    "Stops of people from some ethnic groups are more likely to lead to a ",
    "formal criminal-justice outcome: the probability of a formal outcome to ",
    "is ",
    to_text(pull(
      arrange(filter(outcomes_by_ethnicity_model, sig), desc(odds_ratio)), 
      diff_type
    )),
    "\n\n[{report_q}_07_chart_outcomes_ethnicity.png]"
  ),
  paste0(
    "The likelihood of searches having a formal criminal-justice outcome is ",
    "similar across ethnicities – over the past 12 months, ",
    ifelse(
      sum(outcomes_by_ethnicity_model$sig) > 0, 
      str_glue(
        "the probability of a formal outcome was similar across ethnic ",
        "groups, except for ",
        to_text(pull(filter(outcomes_by_ethnicity_model, sig), search_type))
      ),
      str_glue(
        "the probability of a formal outcome to searches of Black or Asian ",
        "people was not significantly different from the probability of a ",
        "formal outcome to searches of White people for any of the main ",
        "search types"
      )
    ),
    str_glue(
      "\n\n[{report_q}_07_chart_outcomes_ethnicity.png]\n\nALT: Bar charts ", 
      "showing the proportion of searches of different types resulting in a ", 
      "formal criminal-justice outcome varies across ethnic groups"
    )
  )
)

tweets[8] <- str_glue(
  "Half of searches from {report_period} occurred in ",
  "{percent(stop_concentration)} of neighbourhoods – there were more searches ",
  "in {searches_by_ward$ward[1]} ward in {searches_by_ward$borough[1]} than ",
  "in {places_less_than_top_ward}.\n\n[{report_q}_08_chart_map.png]\n\nALT: ",
  "Map showing where stop and searches are concentrated in London"
)

tweets[9] <- str_glue(
  "The full report – created in #RStats using #OpenData – is available at ", 
  "http://lesscrime.info/publication/stop-and-search-london-",
  "{str_replace(report_q, '_', '-')}/\n\nWe update the report every quarter ",
  "to provide the latest analysis of stop and search in London\n\n",
  "[{report_q}_00_cover.png]\n\nALT: Cover image for stop and search report\n"
)

tweets %>% 
  paste(collapse = "\n\n-----\n\n") %>% 
  write_file(here::here(str_glue("output/{report_q}/{report_q}_tweets.md")))
```



\restoregeometry

```{r chart weekly-hit-rate}
weekly_hit_rate <- stops %>% 
  filter(object %in% main_types) %>% 
  mutate(
    object = case_when(
      object == "weapons" & 
        legislation == "Police and Criminal Evidence Act 1984 (section 1)" ~ 
        "weapons (based on reasonable suspicion)",
      object == "weapons" & 
        legislation == "Criminal Justice and Public Order Act 1994 (section 60)" ~ 
        "weapons (based on s.60 authorisation)",
      object == "weapons" ~ "unknown",
      TRUE ~ object
    )
  ) %>% 
  filter(object != "unknown") %>% 
  count(
    week = yearweek(date), 
    positive = !outcome %in% c("no further action", "unknown"),
    object
  ) %>% 
  group_by(week, object) %>% 
  mutate(
    searches = sum(n), 
    hit_rate = n / sum(n)
  ) %>% 
  filter(positive) %>% 
  group_by(object) %>% 
  filter(
    between(searches, quantile(searches, 0.025), quantile(searches, 0.975))
  ) %>%
  ungroup()

chart_weekly_hit_rate <- ggplot(weekly_hit_rate, aes(searches, hit_rate)) +
  geom_smooth(
    aes(colour = "colour"), 
    method = "loess", 
    formula = "y ~ x", 
    # se = FALSE,
    span = 2
  ) +
  geom_point(aes(fill = "fill"), alpha = 0.5, size = 0.2) +
  facet_wrap(
    vars(object), 
    scales = "free",
    labeller = as_labeller(function (string) {
      str_wrap(paste("searches for", string), 30)
    })
  ) +
  scale_x_continuous(
    labels = comma_format(accuracy = 1), 
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.25)
  ) +
  scale_colour_manual(
    values = c("colour" = unname(ucl_colours_list["Bright Blue"])),
    labels = c("colour" = "relationship between\nsearches and hit rate")
  ) +
  scale_fill_manual(
    values = c("fill" = "black"), 
    labels = c("fill" = str_glue(
      "weekly count of\nstop and search in London,\n{format(min(stops$date), '%B %Y')} to ",
      "{format(max(stops$date), '%B %Y')}"
    ))
  ) +
  labs(
    title = "**Increasing the frequency of stop and search in London has diminshing returns**",
    subtitle = str_glue(
      "For most search types, the proportion that result in a formal outcome ",
      "(such as an arrest) goes down as the number goes up. One <br>",
      "exception is firearms searches, for which the rate of formal outcomes ",
      "stays the same as the number of searches changes."
    ),
    caption = str_glue(
      "Note: outliers in the top and bottom 2.5% of weekly search counts ",
      "removed to avoid outliers distoring the relationships between variables"
    ),
    x = "searches per week",
    y = "proportion resulting in a formal outcome",
    colour = NULL,
    fill = NULL
  ) +
  theme_stop_search() +
  theme(
    axis.title = element_text(hjust = 0.5),
    legend.margin = margin(0, 0, 0, 0),
    legend.justification = c(1, 0),
    legend.position = c(1, 0),
    plot.caption = element_text(hjust = 0),
    plot.caption.position = "plot",
    plot.title = ggtext::element_markdown(),
    plot.subtitle = ggtext::element_markdown(face = "plain", size = 10, lineheight = 1.2),
    strip.placement = "outside",
    strip.text = element_text(face = "bold")
  )

save_chart(
  chart_weekly_hit_rate, 
  "weekly_hit_rate_by_type", 
  "",
  height = 500, 
  width = 600, 
  title_as_subtitle = FALSE, 
  set_caption = FALSE
)
```

